A “torn bag mechanism” of small extracellular vesicle release via limiting membrane rupture of en bloc released amphisomes (amphiectosomes)

Version 1

Abstract

Abstract

Recent studies showed an unexpected complexity of extracellular vesicle (EV) biogenesis pathways. We previously found evidence that human colorectal cancer cells in vivo release large multivesicular body-like structures en bloc. Here, we tested whether this large extracellular vesicle type is unique to colorectal cancer cells. We found that all cell types we studied (including different cell lines and cells in their original tissue environment) released multivesicular large EVs. We also demonstrated that upon spontaneous rupture of the limiting membrane of the multivesicular large EVs, their intraluminal vesicles (ILVs) escaped to the extracellular environment by a “torn bag mechanism”. We proved that the multivesicular large EVs were released by ectocytosis of amphisomes (hence, we termed them amphiectosomes). Both ILVs of amphiectosomes and small EVs separated from conditioned media were either exclusively CD63 or LC3B positive. According to our model, upon fusion of multivesicular bodies with autophagosomes, fragments of the autophagosomal inner membrane curl up to form LC3B positive ILVs of amphisomes, while CD63 positive small EVs are of multivesicular body origin. Our data suggest a novel common release mechanism for small EVs, distinct from the exocytosis of multivesicular bodies or amphisomes, as well as the small ectosome release pathway.

Introduction

Extracellular vesicles (EVs) are phospholipid bilayer enclosed structures1–3, which have important roles in cellular homeostasis and intercellular communication. Exosomes have been defined as small (∼ 50-200 nm) EVs (sEVs) of endosomal origin1, 3. Although autophagy is a major cellular homeostatic mechanism, and is implicated in a broad spectrum of human diseases, the intersection of autophagy and exosome secretion remains poorly understood. Recently, regulatory interactions have been shown between autophagy-related molecules and EV biogenesis4, 5. Furthermore, the LC3-conjugation machinery was demonstrated to specify the cargo packaged into EVs6. Importantly, both others and we reported the secretion of LC3-carrying exosomes6, 7. Particularly relevant to the findings presented here, is the implication of amphisomes (hybrid organelles formed by the fusion of late endosomes/multivesicular bodies (MVBs) with autophagosomes8, 9) in EV biogenesis. It was suggested that fusion of the limiting membrane of amphisomes with the plasma membrane of cells results in a subsequent release of exosomes by exocytosis1, 3, 10. The current study was prompted by our recent data showing the in vivo en bloc release of large, MVB-like sEV clusters by human colorectal cancer cells11. Here we investigated if this was a colorectal cancer cell-specific phenomenon. Unexpectedly, we found that it was a general, autophagy-related mechanism of sEV release that we designated as “torn bag mechanism”.

Results and Discussion

In this study, we analyzed in situ fixed, cultured cells with the released EVs preserved in their original microenvironment on a gelatin/fibronectin coated surface, and we detected large multivesicular EVs (MV-lEVs) in sections of different immersion fixed organs. We tested tumorous HT29, HepG2, and non-tumorous HEK293, HEK293T-PalmGFP, HL1 cell lines, as well as primary suspension type bone marrow derived mast cells (BMMCs). In addition, we studied ultrathin sections of mouse kidney and liver.

By the analysis of transmission electron micrographs of all tested cell types, we identified budding (Fig.1A-G) and secretion (Fig.1H-N) of MV-lEVs carrying intraluminal vesicles (ILVs). Importantly, in all cases we found evidence for the extracellular rupture of the limiting membrane of MV-lEVs and the release of ILVs (Fig.1O-U). For this novel type of sEV release, we suggest the designation “torn bag mechanism” which is distinct from the exocytosis of MVBs and amphisomes1, 3, 10 and from the release of plasma membrane derived small EV (sEVs) by ectocytosis12.

Transmission electron microscopic detection of the release and extracellular fate of large, multivesicular EVs secreted by different cell lines and cells in mouse organs

Major steps of the release of large, multivesicular EVs were detected in the case of all tested cell lines including the immortal, non-tumorous HEK293T-PalmGFP (A,H,O), HEK293 (B,I,P), the tumorous cell lines HepG2 (C,J,Q) and HT29 (D,K,R), the beating cardiomyocyte cell line HL1 (E,L,S) and the primary suspension of bone marrow derived mast cells (BMMCs) (F,M,T). The different phases of EV secretion were also captured in the circulation of mouse kidney (V) and liver (W,X). According to the electron micrographs, we found evidence for the budding (A-G,X) and secretion (H-N,V,W) of the multivesicular large EVs (MV-lEVs). We also detected the extracellular rupture of the limiting membrane of the released MV-lEVs with the escape of the intraluminal vesicles (ILVs) by a “torn bag mechanism” (O-U,V). Although it is not always clear whether the secreted MV-lEVs have a single or double limiting membrane, several micrographs suggest the presence of the double membrane (Y-AF) in the secreted MV-lEVs. In the case of BMMCs (Y), the release phase of a multivesicular structure is captured. The bottom portion of this structure embedded in the cytoplasm is surrounded by a single membrane (white arrow head) while the upper (budding) portion is covered by double membrane (asterisk).

Most relevant to the in vivo conditions, we also observed the same phenomenon within the blood vessels of ultrathin sections of both murine kidney (Fig.1V) and liver (Fig.1W,X). In these cases, both the intact MV-lEVs (Fig.1V-X) and the “torn bag release” of sEVs (Fig.1V) were detected. Fig.1X shows that a circulating leukocyte releases MV-lEVs by ectocytosis.

Next, we decided to further investigate the subcellular origin of the ILVs within the secreted MV-lEVs. We analyzed the microenvironment of in situ fixed HEK293T-PalmGFP cells by confocal microscopy. The PalmGFP signal of HEK293T-PalmGFP cells principally associates with the plasma membrane13 (Fig.2_S1C-E). The green fluorescence helped us to identify the plasma membrane-derived limiting membrane of MV-lEVs. In agreement with our previous findings on HT29 colorectal cancer cells, within the MV-lEVs, we found CD63/ALIX (Fig.2A,G) CD81/ALIX (Fig.2B,H), CD63/TSG101 (Fig.2.C,I) and CD81/TSG101 (Fig.2D,J) double positive ILVs or ILV clusters. In line with a recent finding12, CD81 was present in the plasma membrane-derived limiting membrane (Fig.2B,D,F), while CD63 was only found inside the MV-lEVs (Fig.2A,C,E).

Detection of conventional sEV markers and the LC3 protein in HEK293T-PalmGFP cell-derived EVs

Widely used sEV markers (CD63, CD81, ALIX and TSG101) and LC3B were tested in MV-lEVs found in the microenvironment of the releasing cells by confocal microscopy after in situ fixation (A-F). Normalized fluorescence intensities were calculated in order to determine the relative localization of the limiting membrane (PalmGFP), the conventional sEV markers and the LC3B signal (G-L). Fluorescence intensity peaks of sEV markers were largely overlapping with each other, while the LC3B signal and the sEV markers showed separation. Co-localization rates were also calculated (M). The sEV markers co-localized with one another as no significant difference was found among them. In contrast, low co-localization rates were detected between the “classical” sEV markers and LC3B (one-way ANOVA, p<0.0001, n=8-26 confocal images). Real time release of LC3 positive sEVs by the “torn bag mechanism” was studied in the case of HEK293T-PalmGFP-LC3RFP cells by Elyra7 SIM2 super-resolution live cell imaging (N,O). Images were recorded continuously, and selected serial time points are shown. LC3 positive, red fluorescent small particles were released within a 5 min timeframe (O) and are indicated by white arrows. Presence of CD63 and LC3B were detected in the case of a separated sEV fraction using immunogold TEM. HEK293T-PalmGFP derived sEV fraction is shown by negative-positive contrast without immune labelling (P). In double-labelled immunogold TEM images (Q,R), distinct LC3B positive (Q) and CD63 positive (R) sEVs were found. However, CD63-LC3B double positive EVs were not detected. Black arrowheads indicate 10 nm gold particles identifying LC3B, while white arrowheads show 5 nm gold particles corresponding to the presence of CD63. Quantitative analysis of TEM images was performed (S), and the diameters of different EV populations were determined. The LC3B negative population was significantly smaller than the LC3B positive one (p<0.0001, t-test; n=79-100). No difference was detected when the immunogold labelled sEV fraction (either LC3B positive or negative, LC3B+/-) and the unlabeled sEV fraction (sEV) were compared (p<0.05, t-test, n=112-179).

Confocal microscopic analysis of HEK293T-PalmGFP and HepG2 cells derived lEVs; HEK293T-PalmGFP cells

The MV-lEVs released by HEK293T-PalmGFP (A) and HepG2 cells (B) were tested for CD63 sEV and Rab7 late endosomal markers. Both CD63 and Rab7 were present in the ILVs. The GFP positivity of the HEK293T-PalmGFP cells was controlled by immunofluorescence microscopy on control (C) and Chloroquine-treated (D) cells. Co-localization of Palm-GFP and the anti-GFP signal was calculated in the plasma membrane and in the cytoplasm (E) and Pearson’s correlations were visualized. The green fluorescence in the plasma membrane was clearly GFP-dependent in the case of control (Ctrl Membrane) and Chloroquine-treated (Chloro Membrane) samples, while in the cytoplasm, the correlation was significantly weaker (Ctrl Cytoplasm and Chloro Cytoplasm). Within the cytoplasm, the correlation was stronger in Chloroquine-treated cells, suggesting that the phagophore membrane may accumulate Palm-GFP (p < 0.0001, t-test; n = 45).

Next we studied the possible autophagy related aspects of the secreted MV-lEVs. ILVs were tested for the autophagy marker LC3B in parallel with CD63 and CD81. Although LC3B, CD63 and CD81 were all present in association with the ILVs (Fig.2E,F), the LC3B and CD63 (Fig.2K) and the LC3B and CD81 (Fig.2L) signals did not overlap. Fig.2M shows that while the known sEV markers (CD63, CD81, TSG101 and ALIX) strongly co-localized with each other, LC3B positivity hardly showed co-localization with CD63 or CD81. Immunocytochemistry results of HT29 and HepG2 cells further validated the findings obtained with the HEK293T-PalmGFP cells (Fig.2_S2). The ILVs of HEK293T-PalmGFP and HepG2 cell lines were Rab7 positive (Fig.2_S1A,B) suggesting a late endosomal origin. Western blot validation of the applied antibodies is summarized in Fig.2_S3. The sEV markers were also tested on separated sEVs of HEK293T-PalmGFP cells by transmission electron microscopy (TEM). Fig.2P confirms the typical sEV morphology. With TEM double immunogold labelling, using anti-LC3B and anti-CD63 antibodies simultaneously, we found distinct LC3B positive (Fig.2Q, Fig.2_S5O) and CD63 positive (Fig.2R, Fig.2_S5O) sEVs. Based on the analysis of TEM images, the diameters of unlabeled, and LC3B positive and negative sEVs were determined (Fig.2S). The LC3B positive sEVs had a significantly larger diameter as compared to the LC3B negative ones. The simultaneous presence of CD63, CD81, TSG101, ALIX and the autophagosome marker LC3B within the MV-lEVs (rather than on their limiting membrane) identified the extracellular MV-lEVs as en bloc released amphisomes.

Confocal microscopic images of amphiectosome release by HT29 and HepG2 cells

The intraluminal EVs of amphiectosomes were found to be positive for LC3B and CD81 in HT29 cultures (A) and HepG2 cells (H). In the case of HT29 cells, phases of the “torn bag mechanism” were captured, including a secreted intact amphiectosome (B), MV-lEV with ruptured limiting membrane releasing internal vesicles (C, white arrow), an inside-out secreted amphiectosome (D, white arrow), and an amphiectosome with a fully disintegrated limiting membrane and released sEVs (E). The presence of TSG101/CD63 (F), ALIX/CD81 (G), and LC3B/CD81 (H) was also examined in the released amphiectosomes of the HepG2 cell line.

Western blot validation of antibodies used in immune-fluorescence detection

Lysates from three distinct human cell lines (HEK293, HepG2 and HT29) were employed in the validation process to reduce the cell line-specific variations in the study. Protein bands lacking posttranslational modifications are noted as “w/o PTM.” Numerous posttranslational modifications of CD63 and CD81 are widely recognized.

Characterization of the HEK293T-PalmGFP-LC3RFP cell line.

Confocal microscopy (A) and Western blot analysis (B) were employed to characterize the HEK293T-PalmGFP-LC3RFP cell line. For overnight Chloroquine treatments, 30 μM Chloroquine was applied. Punctate LC3 fluorescence observed in (A) corresponds to autophagosomes. In panel (B), whole-cell lysates from HEK293T-PalmGFP (lane i), HEK293T-PalmGFP-LC3RFP (lane ii) and Chloroquine-treated HEK293T-PalmGFP-LC3RFP (lane iii) samples were examined. Gray arrows indicate LC3I, while black arrows highlight LC3II. Full Western Blot images can be find in Fig.3_S2 supplementary figure.

Structures involved in amphiectosome release

Multivesicular body (MVB, A), autophagosome (B), amphisome (C), amphiectosome (D) and secreted sEVs (E) were identified by TEM with and without immunogold labelling of HEK293T-PalmGFP cell cultures. White arrowheads (5 nm gold particles) indicate CD63 and black arrowheads (10 nm gold particles) show LC3B. While MVBs (F) were LC3B negative (K), we detected CD63 positivity on the surface of the ILVs (K). In an autophagosome (G), the limiting membrane layers were double positive for CD63 and LC3B (L). In contrast, the internal membranes of autophagosome were CD63 single positive (L). In the case of an amphisome (H), heterogeneous membrane structures were visible with variable size and morphology. The ILVs were either CD63 or LC3B positive (M). The amphiectosomes were located in the extracellular space and contained ILVs of different size and shape (I). The ILVs of amphiectosome (as in case of amphisome), were either CD63 or LC3B positive (N). Secreted sEVs (J) with immunogold labelling (O), were also found to be either CD63 or LC3B positive. Release of an amphiectosome is shown (P) with CD63 and LC3B immunogold signals. Higher magnification of the insert is indicated by the black rectangle. It shows either CD63 or LC3B positive ILVs. Size distributions of ILVs of MVBs, amphisomes and amphiectosomes were determined on Epon-embedded ultrathin sections (Q). Although the ILV sizes differed significantly (one-way ANOVA, ****: p<0.0001, n=73, 138 and 595, respectively), the majority of ILVs had a diameter between 40-100 nm. The diameter of LC3B positive and negative ILVs of amphiectosomes was assessed on TEM images of immunogold labelled ultrathin sections (R). LC3B negative ILVs were significantly smaller than the LC3B positive ones, while the ILVs in the Epon-embedded sections did not differ from the LC3B positive ones (one-way ANOVA, p <0.001, n= 595, 101 and 70, respectively).

The “torn bag mechanism” was also monitored by live cell SIM2 super-resolution microscopy analysis of HEK293T-PalmGFP-LC3RFP cells (Fig.2N,O). The release of the LC3 positive red fluorescent signal was detected within a relatively short period of time (260 sec) (Fig.2O). We could rule out the possibility that rupture of the limiting membrane detected by TEM (Fig.1O-T,V) was a fixation artefact by showing the release of LC3 positive sEVs from amphiectosomes with live-cell imaging. Characterization of the in-house developed HEK293T-PalmGFP-LC3RFP cell line is shown in Fig.2_S4.

In the following step, we addressed the question whether LC3, associated with the ILVs of MV-lEVs, indeed reflected autophagy origin. We tested MVBs (Fig.2_S5A,F,K), autophagosomes (Fig.2_S5B,G,L), amphisomes (Fig.2_S5C,H,M), amphiectosomes (Fig.2_S5D,I,N) and isolated sEV fractions of the same cells (Fig.2_S5E,J,O). Using immune electron microscopy, as expected, we found CD63 single positivity in MVBs (Fig.2_S5K). In autophagosomes, the limiting phagophore membrane was LC3B positive, and CD63 positivity was also present (Fig.2_S5L). The limiting membrane of amphisomes was LC3B negative, and the internal membranous structures were either LC3B or CD63 positive (Fig.2_S5M). The same immunostaining was also observed in the ILVs of the released amphiectosomes (Fig.2_S5N). Importantly, the separated sEVs of HEK293T-PalmGFP cells were either LC3B or CD63 positive (Fig.2_S5O). Thus, we confirmed our confocal microscopy results at the ultrastructural level. Using immunogold TEM, we provided further evidence for the budding/ectocytosis mechanism of amphiectosome release (Fig.2_S5P). The diameters of ILVs within MVBs, amphisomes and amphiectosomes were compared (Fig.2_S5Q), and the differences were likely due to the different pH and osmotic conditions within these organelles. In agreement with our observations with separated sEVs, LC3B positive ILVs had a significantly larger diameter than the LC3B negative ones (Fig.2_S5R) possibly indicating their different intracellular origin. Based on all the above findings, we propose the following model (Fig.3A): autophagosomes and MVBs fuse to form amphisomes, and the inner, LC3 positive membrane of autophagosomes disintegrates14. It curls up and forms LC3 positive ILVs. Therefore, amphisomes contain both MVB-derived and autophagosome-derived, LC3 negative and positive ILVs, respectively. The amphisome is next released from the cell by ectocytosis. Finally, the plasma membrane-derived limiting membrane ruptures enabling the ILVs escape to the extracellular space by a “torn bag mechanism”.

Amphiectosome release and its modulation

Based on our data, a model of amphiectosome release was generated (A). According to this model, the fusion of MVBs and autophagosomes forms amphisomes. The LC3B positive membrane layer (indicated in orange) undergoes disintegration and forms LC3B positive ILVs inside the amphisome. Later, the amphisome is released into the extracellular space by ectocytosis and can be identified extracellularly as an amphiectosome. Finally, the limiting membrane(s) of the amphiectosome is ruptured and the ILVs are released as sEVs into the extracellular space by a “torn bag mechanism”. In order to further support our model on amphiectosome release and “torn bag” EV secretion different in vitro treatments were applied. Cytochalasin B, Colchicine, Chloroquine, Bafilomycin A1 and Rapamycin were used to modulate amphisome release. Points of actions are summarized (B). While Cytochalasin B inhibits actin-dependent membrane budding and cell migration, Colchicine blocks the microtubule-dependent intracellular trafficking. Chloroquine and Bafilomycin have similar, Rapamycin have opposite effects on lysosome-autophagosome or lysosome-amphisome fusion. Chloroquine and Bafilomycin inhibits lysosomal degradation while Rapamycin accelerates it. Based on confocal microscopy, Cytochalasin B (CytoB) did not alter the dynamics of amphiectosome release (C). In contrast, both Colchicine (Colch) and Rapamycin (Rapa) significantly inhibited the release of amphiectosomes, while Chloroquine (Chloro) and Bafilomycin (Bafilo) increased the release frequency. Between Chloroquine and Bafilomycin effect significant difference was not detected (C). Results are shown as mean ± SD of 3-4 independent biological replicates, analyzed by one-way ANOVA, *: p<0.05, **: p<0.01, ns: non-significant. Presence of membrane-bound (lipidated) LC3II was tested by Western blotting. The total protein content of serum-, cell- and large EV-depleted conditioned medium of HEK293T-PalmGFP (PalmGFP) and HEK293T-PalmGFP-LC3RFP (PalmGFP-LC3RFP) cells was purified by precipitation and 20 µg of the protein samples were loaded on the gel (D). The lipidated LC3II band was detected in all cases. Relative expression of control (Ctrl) and Chloro treated samples were determined by densitometry. Chloroquine treatment increased the LC3II level by approximately by two fold. Results are shown as mean ± SD of n=6 biological replicates.

To investigate the process of amphiectosome release, we exposed the MV-lEV releasing cells to different in vitro treatments (Fig.3B). The release of MV-lEVs was monitored by confocal microscopy of in situ fixed cell cultures. Optimal test conditions were determined (Fig.3_S1A-F) and the results are summarized in Fig.3C. Cytochalasin B did not have any effect on the discharge of MV-lEVs suggesting that the release did not involve a major actin-dependent mechanism. In contrast, there was a significant reduction of the MV-lEV secretion upon exposure of the cells to Colchicine indicating a role of microtubules in the release of the MV-lEVs. While Rapamycin significantly reduced the discharge of MV-lEVs, Chloroquine and Bafilomycin induced an enhanced MV-lEV secretion. Rapamycin activates autophagic degradation15, therefore, it induces a shift towards degradation as opposed to secretion. The lysosomotropic agents Chloroquine and Bafilomycin are known to interfere with the acidification of lysosomes16, 17. By blocking the degradation pathway of MVBs/amphisomes (Fig.3B), an enhanced sEV secretion is observed. This effect is well known for exosome secretion from MVBs18, 19. The diameters of the released MV-lEVs were determined based on confocal images (Fig.3_S1G). Metabolic activity of the cells was determined by a Resazurin assay, and a significant reduction was detected upon exposure of the cells to Rapamycin (Fig.3_S1D) in line with previously published data20. LC3II is the membrane-associated, lipidated autophagic form of LC321 and it is the hallmark of autophagy related membranes14. Importantly, by Western blot, we not only showed the presence of the membrane-bound LC3II in serum-free, lEV-depleted (sEV containing) conditioned medium of both HEK293T-PalmGFP and HEK293T-PalmGFP-LC3RFP cells, but the amount of LC3II substantially increased upon Chloroquine treatment (Fig.3D). Raw data of Western blots are available in Fig.3_S2.

Dose determination for treatments and size distribution of MV-lEVs

Relative metabolic activity was assessed through a Resazurin assay (A-E) during treatment optimization. The red dashed line indicates 100 % metabolic activity, representing control cells. Results are presented as mean ± SD values of n=3-4 independent biological replicates. Student’s unpaired t-test was performed to compare control and treated cells (*: p<0.05, **: p<0.01, ****: p<0.0001). For Colchicine treatment, alterations in the microtubular network were observed through immunocytochemistry (F) and documented using an epifluorescent microscope. Changes in the diameter of released amphiectosomes under various treatments were determined on confocal microscopy images. A significant reduction in size was identified only in the case of Rapamycin treatment (G, one-way ANOVA test *: p<0.05, n= 95-101).

Unedited Western blots of Fig.2_S3 and Fig.3

Subfigure A displays the unedited Western blots utilized for the quantification in Fig.3D. Black arrows indicate the location of LC3I, while red squares represent the quantified LC3II. In Subfigure B, the unedited blots from Fig.2_S3B are presented. Black arrows denote the LC3I-RFP fusion protein, while red arrows indicate the LC3II-RFP fusion protein. The blue arrow points to the native LC3I, while the green arrow represents the native LC3II protein.

Recent advances in the EV field shed light on migrasomes, a special type of MV-lEVs22, 23. With their pomegranate-like ultrastructure, migrasomes resemble amphiectosomes. Therefore, we tested the presence of TSPAN4, a migrasome limiting membrane marker23, in amphiectosomes. Fig.4A,B,G,H show that although TSPAN4 was present intraluminally in the HEK293T-PalmGFP-derived MV-lEVs, it was clearly absent from their limiting membrane. Surprisingly, we identified two different HT29 cell-derived MV-lEV populations: one with intraluminal TSPAN4 only (Fig.4C,E,I,K), and another one with a TSPAN4 positive limiting membrane (Fig.4D,F,J,L). This raised the possibility that the latter population corresponded to migrasomes. Our co-localization analysis also confirmed the existence of two distinct MV-lEV populations (Fig.4M). Next, we carried out live-cell imaging on HEK293T-PalmGFP-LC3RFP cells. The released MV-lEVs were either LC3 positive or negative intraluminally (Fig.4N,O). Our TEM images confirmed that certain cell types can release both migrasome-like structures and amphiectosomes. MV-lEVs with typical migrasome-associated retraction fiber(s) were detected in the case of HL1 (Fig.4P), HEK293T-PalmGFP (Fig.4Q) and BMMC cells (Fig.4R). Importantly, the same cells also released amphiectosomes by budding from the cell surface (Fig.4S-U). Taken together, based on the absence of TSPAN4 in their limiting membrane and their lack of association with retraction fibers, amphiectosomes appear to be distinct from migrasomes. Besides migrasomes, another MV-lEV type was described in the case of gastrointestinal tumors and low-grade glioblastoma cells referred to as spheresome24, 25. However, there is no data on a relationship of spheresome release and autophagy. Recently, endothelial cell derived, multicompartmented microvesicles (MCMVs) were shown to protrude and pinch-off from the cell surface releasing ILVs by a mechanism similar to exocytosis26. The absence of protrusion clusters described for MCMV26 distinguishes amphiectosomes from MCMVs.

Comparison of amphiectosomes and migrasomes

Commonly used sEV markers (CD63, CD81) and TSPAN4, a suggested migrasome marker, were tested in in situ fixed intact MV-lEVs of HEK293T-PalmGFP (A,B) and HT29 (C-F) cells by confocal microscopy. Normalized fluorescence intensities were calculated in order to determine the relative localization of the limiting membrane (with PalmGFP or lactadherin labelling) and the CD63/TSPAN4 and CD81/TSPAN4 markers (G-L). In the case of HEK293T-PalmGFP-derived EVs, there were no “classical migrasomes” with TSPAN4 in their limiting membrane were found. The TSPAN4 signal was only detected intraluminally in the MV-lEVs. The limiting membranes of HT29-derived MV-lEVs were either TSPAN4 positive and negative. The co-localization rate between the limiting membrane and TSPAN4 was low in case of HEK293T-PalmGFP-derived EVs. In the case of HT29 cells, two MV-lEV populations were identified: one with low and one with high co-localization rates (M). Live cell imaging of HEK293T-PalmGFP-LC3RFP cells showed migrasomes-like (retraction fiber-associated) MV-lEVs with or without intraluminal LC3 positivity (N,O). Using TEM, we could identify structures with retraction fiber-associated migrasome morphology in the case of HL1 (P), HEK293T-PalmGFP (Q) and BMMC (R) cells. For comparison, budding of amphiectosomes of the same HL1 (S), HEK293T-PalmGFP (T) and BMMC cells (U) are shown (without being associated with long retractions fibers).

Our approach, involving in situ fixation of cultures and tissues, made it possible to recognize sEV release from amphiectosomes by the “torn bag mechanism”. We propose that this mechanism can be easily missed if conditioned medium is subjected to centrifugation, SEC purification or even to simple pipetting, which may rupture of the limiting membrane of amphiectosomes. This aligns with our observation that the spontaneous escape of ILVs from untouched amphiectosomes can occur as early as 5 minutes after amphiectosome release. Our data presented here suggest that amphiectosome secretion and the “torn bag mechanism” may have a significant, however, previously unrecognized role in sEV biogenesis.

Material and methods

Cell lines

The HEK293 human embryonic kidney, the HepG2 human hepatocyte carcinoma cell line and the HT29 human colon adenocarcinoma cell lines were purchased from the European Collection of Authenticated Cell Cultures (ECACC) through their distributor (Sigma). The HL1 cell line was purchased from Millipore. The HEK293T-PalmGFP human embryonic kidney cells was kindly provided by Charles P. Lai27. Mouse bone marrow-derived mast cells (BMMCs) were differentiated and expanded as we described previously28. The HEK293, HEK293T-PalmGFP and HepG2 cell lines were grown in DMEM (Gibco)29, 30, the HT29 cells were cultured in RPMI 1640 (Gibco)11, while the HL1 cells were grown in Claycomb medium31. All cells were cultured with 10 % fetal bovine serum (FBS, BioSera) in the presence of 100 U/mL of Penicillin and 100 µg/mL Streptomycin (Sigma).

Before analysis by confocal microscopy, the cells were cultured on the surface of gelatin-fibronectin coated glass coverslips (VWR). The coating solution contained 0.02 % gelatin (Sigma) and 5 mg/mL fibronectin (Invitrogen). Coverslips were coated overnight (O/N) at 37 °C.

For transmission electron microscopy (TEM), the adherent cells (HEK293, HEK293T-PalmGFP, HepG2, HT29 and HL1) were grown in gelatin-fibronectin coated 8-well Flux Cell Culture Slides (SPL).

Cell cultures were tested regularly for Mycoplasma infection by PCR, with the following PCR primers: GAAGAWATGCCWTATTTAGAAGATGG and CCRTTTTGACTYTTWCCAC-CMAGTGGTTGTTG 31.

Generation of HEK293T-PalmGFP-LC3RFP cell line

For generation of a stable HEK293T-PalmGFP-LC3RFP cell line, HEK293T-PalmGFP cells were transfected by LentiBrite RFP-LC3 Lentiviral particles (Merck) according to the instructions of the manufacturer. The GFP-RFP double positive cells were sorted by a HS800 Cell Sorter (SONY) and cell banks (master and working cell banks) were prepared. The success of the stable transfection was analyzed by immunocytochemistry and Western blotting. Results are shown in Fig.2_S3.

Confocal microscopy

Confocal microscopy was carried out as we described earlier28, 31 with some modifications. Briefly, unlike the majority of the studies in the field of EVs, here we analyzed untouched, in situ fixed and cultured cells and their microenvironment. Since centrifugation may disrupt the limiting membrane of amphiectosomes, the in situ fixation made it possible to observe them in their intact form. The culture medium was gently removed by pipetting from above the cells leaving a thin medium layer only (approximately 150 µL of liquid on the cells). Without any further washing, cells were in situ fixed by 4 % paraformaldehyde (PFA) in phosphate buffer saline (PBS) for 20 min at room temperature (RT). The released lEVs were either fixed and captured during the release or were preserved on the gelatin/fibronectin surface coating. After fixation, 3x 5 min washes with 50 mM glycine in PBS were carried out. In the case of the non-fluorescent HepG2 and HT29 cells, a lactadherin-based plasma membrane staining was performed13, 28, 31. Lactadherin (Haematologic Technologies) was conjugated to ATTO488 fluorophore (abcam) according to the instructions of the manufacturer. The lactadherin-ATTO488 conjugate was added to the cells in 1:100 dilution in PBS (for 1 h, RT). The unbound lactadherin was removed by washing with PBS (3 times, 5 min, RT) and post-fixation was carried out by 4 % PFA (20 min, RT). PFA was removed by washes with 50 mM glycine in PBS (3 times, 5 min, RT). Blocking and permeabilization of the cells were performed by 10 % FBS with 0.1 % TritonX-100 (Sigma) in PBS (1 h, RT). In general, primary antibodies were applied in 1:200 dilution O/N at 4 °C in the above blocking and permeabilization solution. Excess primary antibodies were eliminated by washing with the blocking and permeabilization solution (3 times, 5 min RT). The secondary antibodies were applied in 1:1000 dilution in 1 % FBS in PBS (1 h, RT). Unbound secondary antibodies were eliminated by washing (1 % FBS, in PBS, 2 times, 5 min; PBS 2 times, 5 min; water 2 times, 5 min) and the samples were mounted in ProLong Diamond with DAPI (Invitrogen). Slides were examined by Leica SP8 Lightning confocal microscope with adaptive lightning mode using a HC PL APO CS2 63x/1.40 OIL objective with hybrid detector. The applied lookup tables (LUT) were linear during this study. For image analysis and co-localization studies, we applied Leica LASX software using the unprocessed raw images. In case of co-localization studies, 20 % threshold and 10 % background settings were applied.

Transmission electron microscopy

Adherent cells (HEK293, HEK293T-PalmGFP, HepG2, HT29 and HL1), as well as mouse (C57BL/6) kidney and liver tissues pieces (approx. 1.5 mm x 1.5 mm) were immersed in and fixed by 4 % glutaraldehyde (48 h/4 °C), post-fixed by 1% osmium tetroxide (2 h, RT) and were embedded into Epon resin (Electron Microscopy Sciences) as described previously32. In the case of BMMCs, 920 μL cell suspension was complemented with 80 μL 50% glutaraldehyde to reach the final 4 % glutaraldehyde concentration. Cells were fixed for 48h at 4°C and were post-fixed by 1% osmium tetroxide (2 h, RT). During sample preparation, BMMCs were collected by gravity-based sedimentation. Due to the high viscosity of Epon resin, BMMCs were embedded in LR White low viscosity resin (SPI Supplies) according to the instructions of the manufacturer. Ultrathin sections (60 nm) were contrasted by uranyl acetate (3.75 %, 10 min, RT) and lead citrate (12 min, RT).

For immunogold TEM studies, cells and tissues were fixed by 4 % PFA with 0.1% glutaraldehyde (48 h, 4 °C) and were post-fixed by 0.5 % osmium tetroxide (30 min, RT). Samples were embedded into LR White and were immunogold labelled as described previously11. The contrast was enhanced by uranyl acetate (3.75 %, 1 min, RT) and lead citrate (2 min, RT). Purified HEK293T-PalmGFP-derived sEV-s were detected by negative-positive contrasting and immunogold labelling as described previously31. Antibodies were used in 1:50 dilution. Detailed list of the used antibodies is available in the Key Resources Table.

For all electron microscopic studies, a JEOL 1011 transmission electron microscope was used. Images were captured with the help of Olympus iTEM software and for image analysis, Image J software was used.

Live cell imaging

The HEK293T-PalmGFP-LC3RFP stable cell line was cultured the same way as HEK293T-PalmGFP cells. Before the experiments, gelatin-fibronectin coated 10-well coverslip bottom chamber slide (Greiner-BioOne) was seeded and treated by 30 μM Chloroquine O/N. Release of migrasomes, amphiectosomes and sEVs were followed by the Leica SP8 Lightning confocal microscope equipped with an Okolab environmental chamber and a Zeiss ELYRA 7 with Lattice SIM² super-resolution fluorescent microscope with the help of 63x/1.4 plan-apochromat Oil objective. For image analysis, we applied Leica LASX, Zeiss ZEN Blue and Image J software.

Modulation of amphiectosome release

In order to test the release mechanism of amphiectosomes and to distinguish them from migrasomes, different treatments were applied overnight (O/N) in fresh, serum containing cell culture medium except for Colchicine, where 1 h treatment was selected. Maturation and fusion of endosomes and lysosomes were inhibited by 15.5 µg/mL (30 μM) Chloroquine (Invitrogen) and 10 nM BafilomycinA1 (Sigma). Actin polymerization was inhibited by 125 ng/mL Cytochalasin B (Sigma). Tubulin polymerization and function were inhibited by 250 pg/mL Colchicine, while an autophagy-related degradation was induced by 50 ng/mL Rapamycin. The selected concentrations were determined based both on literature data and our preliminary experiments (Fig.3_S1). Cellular metabolic activity was determined by a metabolic activity-based Resazurin assay31. Fresh cell culture medium was added to control cultures a day before the in situ fixation. Reagents were diluted in fresh cell culture medium. Leica TCS SP8 Lightning confocal microscope was used for detection of amphiectosome release. A few hundred µm2 sized area with 15-20 µm in height was tile-scanned with a few hundred cells. The MV-lEVs were recognized as CD63 positive EVs surrounded by GFP positive membrane. They were counted and were normalized to the number of nuclei.

Western blotting

Presence of proteins and specificity of the used primary antibodies were confirmed by Western blotting as described previously31. For accurate quantification (free of variations potentially caused by EV purification), we analyzed cell-, serum- and lEV-free conditioned medium. The cells were cultured O/N in a serum-free culture medium. After harvesting, the cells were eliminated by centrifugation (300 g, 10 min at 4 °C) followed by a 2.000g centrifugation (30 min at 4 °C) to eliminate lEVs. Total protein content of the conditioned, serum-, cell- and lEV-free medium was precipitated by trichloroacetic acid as described previously31, 33. The protein pellets were suspended in in cOmplete Protease Inhibitor Cocktail (Roche) containing radio-immunoprecipitation assay (RIPA) buffer.

When whole cell lysate was tested for validation of antibodies and the HEK293T-PalmGFP-LC3RFP cell line, cells were lysed in cOmplete Protease Inhibitor Cocktail (Roche) containing RIPA buffer.

Polyacrylamide gel electrophoresis was carried out using 10 % gels (acrylamide/bis-acrylamide ratio 37.5:1) or any kDa precast gels (Biorad) and a MiniProtean (BioRad) gel running system. For better solubilization of membrane proteins, equal volumes of 0.1 % TritonX-100, Laemmli buffer and samples were mixed as described previously34. Approximately 10-30 µg protein were loaded into each well. Following electrophoretic separation, proteins were transferred to PVDF membranes (Serva). Membranes were blocked with 5 % skimmed milk powder or 5 % BSA in washing buffer for 1 h. Primary antibodies were applied in 1:1000 dilution with the exception of the anti-CD63 (rabbit), anti-CD81 (rabbit) and anti-CD81 (mouse) antibodies where 1:500, 1:2500 and 1:100 dilutions were used, respectively. Peroxidase-labelled secondary antibodies were applied in 1:10 000 dilution. The signals were detected by ECL Western Blotting Substrate (Thermo Scientific) with an Imager CHEMI Premium (VWR) image analyzer system. In case of quantification, equal protein amounts were loaded to the gels. Within a biological replicate, the control and Chloroquine-treated samples were run on the same gels. To enable comparison, the relative expression of control and Chloroquine-treated samples were determined and compared.

Software and statistical analysis

For image capturing, analysis and co-localization studies, Leica LAS X, Zeiss ZEN Blue, Olympus iTEM and Image J software were used. Figures and graphs were generated using GraphPad Prism 9.4.1 and Biorender (BioRender.com). For statistical analysis, standard deviation was calculated. Unpaired two tailed student t-tests and one-way ANOVA were used (* p<0.05, ** p<0.01, *** p<0.001, **** p<0.0001).

eLife Assessment

The authors present evidence that small extracellular vesicles can be secreted from cells inside larger vesicles that they call amphiectosomes, which then tear to release their small vesicle contents. There are questions and concerns relating to the quality of the data and the in vivo significance of the observations. The findings are potentially important but the data are incomplete and the claims are only partially supported.

Reviewer #1 (Public Review):

Summary:

The authors' research group had previously demonstrated the release of large multivesicular body-like structures by human colorectal cancer cells. This manuscript expands on their findings, revealing that this phenomenon is not exclusive to colorectal cancer cells but is also observed in various other cell types, including different cultured cell lines, as well as cells in the mouse kidney and liver. Furthermore, the authors argue that these large multivesicular body-like structures originate from intracellular amphisomes, which they term "amphiectosomes." These amphiectosomes release their intraluminal vesicles (ILVs) through a "torn-bag mechanism." Finally, the authors demonstrate that the ILVs of amphiectosomes are either LC3B positive or CD63 positive. This distinction implies that the ILVs either originate from amphisomes or multivesicular bodies, respectively.

Strengths:

The manuscript reports a potential origin of extracellular vesicle (EV) biogenesis. The reported observations are intriguing.

Weaknesses:

It is essential to note that the manuscript has issues with experimental designs and lacks consistency in the presented data. Here is a list of the major concerns:

(1) The authors culture the cells in the presence of fetal bovine serum (FBS) in the culture medium. Given that FBS contains a substantial amount of EVs, this raises a significant issue, as it becomes challenging to differentiate between EVs derived from FBS and those released by the cells. This concern extends to all transmission electron microscopy (TEM) images (Figure 1, 2P-S, S5, Figure 4 P-U) and the quantification of EV numbers in Figure 3. The authors need to use an FBS-free cell culture medium.

(2) The data presented in Figure 2 is not convincingly supportive of the authors' conclusion. The authors argue that "...CD81 was present in the plasma membrane-derived limiting membrane (Figures 2B, D, F), while CD63 was only found inside the MV-lEVs (Fig. 2A, C, E)." However, in Figure 2G, there is an observable CD63 signal in the limiting membrane (overlapping with the green signals), and in Figure 2J, CD81 also exhibits overlap with MV-IEVs.

(3) Following up on the previous concern, the authors argue that CD81 and CD63 are exclusively located on the limiting membrane and MV-IEVs, respectively (Figure 2-A-M). However, in lines 104-106, the authors conclude that "The simultaneous presence of CD63, CD81, TSG101, ALIX, and the autophagosome marker LC3B within the MV-lEVs..." This statement indicates that CD63 and CD81 co-localize to the MV-IEVs. The authors need to address this apparent discrepancy and provide an explanation.

(4) The specificity of the antibodies used in Figure 2 should be validated through knockout or knockdown experiments. Several of the antibodies used in this figure detect multiple bands on western blots, raising doubts about their specificity. Verification through additional experimental approaches is essential to ensure the reliability and accuracy of all the immunostaining data in this manuscript.

(5) In Figures 2P-R, the morphology of the MV-IEVs does not resemble those shown in Figures 1-A, H, and D, indicating a notable inconsistency in the data.

(6) There are no loading controls provided for any of the western blot data. Additionally, for Figures 2-S4B, the authors should run the samples from lanes i-iii in a single gel.

(7) In Figure 2-S4, is there co-localization observed between LC3RFP (LC3A?) with other MV-IFV markers? How about LC3B? Does LC3B co-localize with other MV-IFV markers?

(8) The TEM images presented in Figure 2-S5, specifically F, G, H, and I, do not closely resemble the images in Figure 2-S5 K, L, M, N, and O. Despite this dissimilarity, the authors argue that these images depict the same structures. The authors should provide an explanation for this observed discrepancy to ensure clarity and consistency in the interpretation of the presented data.

(9) For Figures 3C and 3-S1, the authors should include the images used for EV quantification. Considering the concern regarding potential contamination introduced by FBS (concern 1), it is advisable for the authors to employ an independent method to identify EVs, thereby confirming the reliability of the data presented in these figures.

(10) Do the amphiectosomes released from other cell types as well as cells in mouse kidneys or liver contain LC3B positive and CD63 positive ILVs?

Reviewer #2 (Public Review):

Summary:

The authors had previously identified that a colorectal cancer cell line generates small extracellular vesicles (sEVs) via a mechanism where a larger intracellular compartment containing these sEVs is secreted from the surface of the cell and then tears to release its contents. Previous studies have suggested that intraluminal vesicles (ILVs) inside endosomal multivesicular bodies and amphisomes can be secreted by the fusion of the compartment with the plasma membrane. The 'torn bag mechanism' considered in this manuscript is distinctly different because it involves initial budding off of a plasma membrane-enclosed compartment (called the amphiectosome in this manuscript, or MV-lEV). The authors successfully set out to investigate whether this mechanism is common to many cell types and to determine some of the subcellular processes involved.

The strengths of the study are:

(1) The high-quality imaging approaches used, seem to show good examples of the proposed mechanism.

(2) They screen several cell lines for these structures, also search for similar structures in vivo, and show the tearing process by real-time imaging.

(3) Regarding the intracellular mechanisms of ILV production, the authors also try to demonstrate the different stages of amphiectosome production and differently labelled ILVs using immuno-EM.

Several of these techniques are technically challenging to do well, and so these are critical strengths of the manuscript.

The weaknesses are:

(1) Most of the analysis is undertaken with cell lines. In fact, all of the analysis involving the assessment of specific proteins associated with amphiectosomes and ILVs are performed in vitro, so it is unclear whether these processes are really mirrored in vivo. The images shown in vivo only demonstrate putative amphiectosomes in the circulation, which is perhaps surprising if they normally have a short half-life and would need to pass through an endothelium to reach the vessel lumen unless they were secreted by the endothelial cells themselves.

(2) The analysis of the intracellular formation of compartments involved in the secretion process (Figure 2_S5) relies on immuno-EM, which is generally less convincing than high-/super-resolution fluorescence microscopy because the immuno-labelling is inevitably very sporadic and patchy. High-quality EM is challenging for many labs (and seems to be done very well here), but high-/super-resolution fluorescence microscopy techniques are more commonly employed, and the study already shows that these techniques should be applicable to studying the intracellular trafficking processes.

(3) One aspect of the mechanism, which needs some consideration, is what happens to the amphisome membrane, once it has budded off inside the amphiectosome. In the fluorescence images, it seems to be disrupted, but presumably, this must happen after separation from the cell to avoid the release of ILVs inside the cell. There is an additional part of Figure 1 (Figure 1Y onwards), which does not seem to be discussed in the text (and should be), that alludes to amphiectosomes often having a double membrane.

(4) The real-time analysis of the amphiectosome tearing mechanism seemed relatively slow to me (over three minutes), and if this has been observed multiple times, it would be helpful to know if this is typical or whether there is considerable variation.

Overall, I think the authors have been successful in identifying amphiectosomes secreted from multiple cell lines and demonstrating that the ILVs inside them have at least two origins (autophagosome membrane and late endosomal multivesicular body) based on the markers that they carry. The analysis of intracellular compartments producing these structures is rather less convincing and it remains unclear what cells release these structures in vivo.

I think there could be a significant impact on the EV field and consequently on our understanding of cell-cell signalling based on these findings. It will flag the importance of investigating the release of amphiectosomes in other studies, and although the authors do not discuss it, the molecular mechanisms involved in this type of 'ectosomal-style' release will be different from multivesicular compartment fusion to the plasma membrane and should be possible to be manipulated independently. Any experiments that demonstrate this would greatly strengthen the manuscript.

In general, the EV field has struggled to link up analysis of the subcellular biology of sEV secretion and the biochemical/physical analysis of the sEVs themselves, so from that perspective, the manuscript provides a novel angle on this problem.

Reviewer #3 (Public Review):

Summary:

In this manuscript, the authors describe a novel mode of release of small extracellular vesicles. These small EVs are released via the rupture of the membrane of so-called amphiectosomes that resemble "morphologically" Multivesicular Bodies.

These structures have been initially described by the authors as released by colorectal cancer cells (https://doi.org/10.1080/20013078.2019.1596668). In this manuscript, they provide experiments that allow us to generalize this process to other cells. In brief, amphiectosomes are likely released by ectocytosis of amphisomes that are formed by the fusion of multivesicular endosomes with autophagosomes. The authors propose that their model puts forward the hypothesis that LC3 positive vesicles are formed by "curling" of the autophagosomal membrane which then gives rise to an organelle where both CD63 and LC3 positive small EVs co-exist and would be released then by a budding mechanism at the cell surface that appears similar to the budding of microvesicles /ectosomes. Very correctly the authors make the distinction from migrasomes because these structures appear very similar in morphology.

Strengths:

The findings are interesting despite that it is unclear what would be the functional relevance of such a process and even how it could be induced. It points to a novel mode of release of extracellular vesicles.

Weaknesses:

This reviewer has comments and concerns concerning the interpretation of the data and the proposed model. In addition, in my opinion, some of the results in particular micrographs and immunoblots (even shown as supplementary data) are not of quality to support the conclusions.

Author response:

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public Review):

Summary:

The authors' research group had previously demonstrated the release of large multivesicular body-like structures by human colorectal cancer cells. This manuscript expands on their findings, revealing that this phenomenon is not exclusive to colorectal cancer cells but is also observed in various other cell types, including different cultured cell lines, as well as cells in the mouse kidney and liver. Furthermore, the authors argue that these large multivesicular body-like structures originate from intracellular amphisomes, which they term "amphiectosomes." These amphiectosomes release their intraluminal vesicles (ILVs) through a "torn-bag mechanism." Finally, the authors demonstrate that the ILVs of amphiectosomes are either LC3B positive or CD63 positive. This distinction implies that the ILVs either originate from amphisomes or multivesicular bodies, respectively.

Strengths:

The manuscript reports a potential origin of extracellular vesicle (EV) biogenesis. The reported observations are intriguing.

Weaknesses:

It is essential to note that the manuscript has issues with experimental designs and lacks consistency in the presented data. Here is a list of the major concerns:

(1) The authors culture the cells in the presence of fetal bovine serum (FBS) in the culture medium. Given that FBS contains a substantial amount of EVs, this raises a significant issue, as it becomes challenging to differentiate between EVs derived from FBS and those released by the cells. This concern extends to all transmission electron microscopy (TEM) images (Figure 1, 2P-S, S5, Figure 4 P-U) and the quantification of EV numbers in Figure 3. The authors need to use an FBS-free cell culture medium.

Although FBS indeed contains bovine EVs, however, the presence of very large multivesicular EVs (amphiectosomes) that our manuscript focuses on has never been observed and reported. For reported size distributions of EVs in FBS, please find a few relevant references below:

PMID: 29410778, PMID: 33532042, PMID: 30940830 and PMID: 37298194

All the above publications show that the number of lEVs > 350-500 nm is negligible in FBS. The average diameter of MV-lEVs (amphiectosomes) described in our manuscript is around 1.00-1.50 micrometer.

Reviewer #1: These papers evaluated the effectiveness of various methods to eliminate EVs from FBS, emphasizing the challenges associated with the presence of EVs in FBS. They also caution against using FBS in EV studies due to these issues. However, I did not find a clear indication regarding the size distributions of EVs in FBS in these papers.

Please provide accurate reference supporting the claim that 'lEVs > 350-500 nm are negligible in FBS.' The papers cited by the authors do not address this specific point.

In the revised manuscript, we addressed the point that due to sterile filtering of FBS, it cannot contain large >0.22 µm EVs

Our response to Reviewer #1 point 2. When we demonstrated the TEM of isolated EVs, we consistently used serum- free conditioned medium (Fig2 P-S, Fig2S5 J, O) as described previously (Németh et al 2021, PMID: 34665280).

Reviewer #1: This is an important point that is not mentioned in the original main text, figure legend or method. Please address.

We agree and we apologize for it. We added this information to the revised manuscript.

Our response to Reviewer #1 point 3. Our TEM images show cells captured in the process of budding and scission of large multivesicular EVs excluding the possibility that these structures could have originated from FBS.

Reviewer #1: These images may also depict the engulfment of EVs in FBS. Hence, it is crucial to utilize EV-free or EV-depleted FBS.

As we mentioned earlier, we added the information to the revised manuscript that sterile filtering of the FBS presumably removed particles >0.22 µm EVs

Our response to Reviewer #1 point 4. In addition, in our confocal analysis, we studied Palm-GFP positive, cell-line derived MV-lEVs. Importantly, in these experiments, FBS-derived EVs are non-fluorescent, therefore, the distinction between GFP positive MV-lEVs and FBS-derived EVs was evident.

Reviewer #1: I agree that these fluorescent-labeled assays conclusively indicate that the MV-lEVs are originating from the cells. However, the images of concerns are the non- fluorescent-labeled images in (Figure 1, 2P-S, S5, Figure 4 P-U and Figure 3). The MV-lEVs may derive from both the cells and FBS.

Please see above our response to points 1-3.

Our response to Reviewer #1 point 5. In addition, culturing cells in FBS-free medium (serum starvation) significantly affects autophagy. Given that in our study, we focused on autophagy related amphiectosome secretion, we intentionally chose to use FBS supplemented medium.

Reviewer #1 If this is a concern, the authors should use EV-depletive FBS.

As we discussed above, sterile filtration of FBS removes particles >0.22 µm. In addition, based on our preliminary experiments, EV-depleted serum may effect cell physiology.

Our response to Reviewer #1 point 6. Even though the authors of this manuscript are not familiar with the technological details how FBS is processed before commercialization, it is reasonable to assume that the samples are subjected to sterile filtration (through a 0.22 micron filter) after which MV-lEVs cannot be present in the commercial FBS samples.

Reviewer #1This is a fair comment that needs to be included in the manuscript.

As you suggested, this comment is now included in the revised manuscript

(2) The data presented in Figure 2 is not convincingly supportive of the authors' conclusion. The authors argue that "...CD81 was present in the plasma membrane-derived limiting membrane (Figures 2B, D, F), while CD63 was only found inside the MV-lEVs (Fig. 2A, C, E)." However, in Figure 2G, there is an observable CD63 signal in the limiting membrane (overlapping with the green signals), and in Figure 2J, CD81 also exhibits overlap with MV-IEVs.

Both CD63 and CD81 are tetraspanins known to be present both in the membrane of sEVs and in the plasma membrane of cells (for references, please see Uniprot subcellular location maps: https://www.uniprot.org/uniprotkb/P08962/entry#subcellular_location https://www.uniprot.org/uniprotkb/P60033/entry#subcellular_location). However, according the feedback of the reviewer, for clarity, we will delete the implicated sentence from the text.

Reviewer #1 Please also justify the statement questioned in (3) as these arguments are interconnected.

We hope you find our above responses to your comment acceptable.

(3) Following up on the previous concern, the authors argue that CD81 and CD63 are exclusively located on the limiting membrane and MV-IEVs, respectively (Figure 2-A-M). However, in lines 104-106, the authors conclude that "The simultaneous presence of CD63, CD81, TSG101, ALIX, and the autophagosome marker LC3B within the MV-lEVs..." This statement indicates that CD63 and CD81 co-localize to the MV-IEVs. The authors need to address this apparent discrepancy and provide an explanation.

There must be a misunderstanding because we did not claim or implicate in the text that “CD81 and CD63 are exclusively located on the limiting membrane and MV-IEVs”. Here we studied co-localization of the above proteins in the case intraluminal vesicles (ILVs). In Fig 2. we did not show any analysis of limiting membrane co-localization.

Reviewer #1 I have indicated that this statement is found in lines 104-106, where the authors argue, 'The simultaneous presence of CD63, CD81, TSG101, ALIX, and the autophagosome marker LC3B within the MV-lEVs...' If the authors acknowledge the inaccuracy of this statement, please provide a justification for this argument.

For clarity, we modified the description of data shown in Fig2 in the revised manuscript.

(4) The specificity of the antibodies used in Figure 2 should be validated through knockout or knockdown experiments. Several of the antibodies used in this figure detect multiple bands on western blots, raising doubts about their specificity. Verification through additional experimental approaches is essential to ensure the reliability and accuracy of all the immunostaining data in this manuscript.

We will consider this suggestion during the revision of the manuscript.

Reviewer #1:Please do so.

We carefully considered the suggestion, but we realized that it was not feasible for us to perform gene silencing in the case of all our used antibodies before resubmission of our revised manuscript. However, we repeated the Western blot for mouse anti-CD81 (Invitrogen MAA5-13548) and replaced the previous Western blot by it in the revised manuscript (Fig.2-S4H)

(5) In Figures 2P-R, the morphology of the MV-IEVs does not resemble those shown in Figures 1-A, H, and D, indicating a notable inconsistency in the data.

EM images in Figure2 P-R show sEVs separated from serum-free conditioned media as opposed to MV-lEVs, which were in situ captured in fixed tissue cultures (Fig1). Therefore, the two EV populations necessarily have different size and structure. Furthermore, Fig. 1 shows images of ultrathin sections while in Figure 2P-R, we used a negative-positive contrasting of intact sEV-s without embedding and sectioning.

(6) There are no loading controls provided for any of the western blot data.

Not even the latest MISEV 2023 guidelines give recommendations for proper loading control for separated EVs in Western blot (MISEV 2023 , DOI: 10.1002/jev2.12404 PMID: 38326288). Here we applied our previously developed method (PMID: 37103858), which in our opinion, is the most reliable approach to be used for sEV Western blotting. For whole cell lysates, we used actin as loading control (Fig3-S2B).

Reviewer #1: The blots referenced here (Fig2-S3; Fig2-S4B; Fig3-S2B) were conducted using total cell lysates, not EV extracts. Only one blot in Fig3-S2B includes an actin control. All remaining blots should incorporate actin controls for consistency.

Fig2-S3 (corresponding to Fig2-S4 in the revised manuscript) only shows reactivity of the used antibodies. This Western blot is not intended to serve as a basis of any quantitative conclusions. Fig2-S4 (corresponding to Fig2-S5 in the revised manuscript) includes the actin control. Fig3-S2B shows the complete membrane, which was cut into 4 pieces, and the immune reactivity of different antibodies was tested. The actin band was included on the anti-LC3B blot. For clarity, we rephrased the figure legend.

Additionally, for Figures 2-S4B, the authors should run the samples from lanes i-iii in a single gel.

Please note that in Figure 2- S4B, we did run a single gel, and the blot was cut into 4 pieces, which were tested by anti-GFP, anti-RFP, anti-LC3A and anti-LC3B antibodies. Full Western blots are shown in Fig.3_S2 B, and lanes “1”, “2” and “3” correspond to “i”, “ii” and “iii” in Fig.2-S4, respectively.

Reviewer #1: In the original Figure 2- S4B, the blots were sectioned into 12 pieces. If lanes "i," "ii," and "iii" were run on the same blot, the authors are advised to eliminate the grids between these lanes.

Grids separating the lanes have been eliminated on Fig.2_S4 (now Fig.2_S5 in the revised manuscript).

(7) In Figure 2-S4, is there co-localization observed between LC3RFP (LC3A?) with other MV-IFV markers? How about LC3B? Does LC3B co-localize with other MV-IFV markers?

In Supplementary Figure 2-S4, we showed successful generation of HEK293T-PalmGFP-LC3RFP cell line. In this case we tested the cells, and not the released MV-lEVs. LC3A co-localized with the RFP signal as expected.

Reviewer #1: Does LC3RFP colocalize with MV-IFV markers in HEK293T-PalmGFP-LC3RFP cell line? This experiment aims to clarify the conclusion made in lines 104-106, where the authors assert that 'The concurrent existence of CD63, CD81, TSG101, ALIX, and the autophagosome marker LC3B within the MV-lEVs...'

In the case of PalmGFP-LC3RFP cells, LC3-RFP is overexpressed. Simultaneous assessment of this overexpressed protein with non-overexpressed, fluorescent antibod-detected molecules proved to be challenging because of spectral overlaps and inappropriate signal-noise ratios. Furthermore, in association with EVs, the number of antibody-detected molecules is substantially lower than in cells. Therefore, even though we tried, we could not successfully perform these experiments.

(8) The TEM images presented in Figure 2-S5, specifically F, G, H, and I, do not closely resemble the images in Figure 2-S5 K, L, M, N, and O. Despite this dissimilarity, the authors argue that these images depict the same structures. The authors should provide an explanation for this observed discrepancy to ensure clarity and consistency in the interpretation of the presented data.

As indicated in Material and Methods, Fig 2-S5 F, G, H and I are conventional TEM images fixed by 4% glutaraldehyde 1% OsO4 2h and embedded into Epon resin with a post contrasting of 3.75% uranyl acetate 10 min and 12 min lead citrate. Samples processed this way have very high structure preservation and better image quality, however, they are not suitable for immune detection. In contrast, Fig.2.-S5 K,L,M,N shows immunogold labelling of in situ fixed samples. In this case we used milder fixation (4% PFA, 0.1% glutaraldehyde, postfixed by 0.5% OsO4 30 min) and LR-White hydrophilic resin embedding. This special resin enables immunogold TEM analysis. The sections were exposed to H2O2 and NaBH4 to render the epitopes accessible in the resin. Because of the different applied techniques, the preservation of the structure is not the same. In the case of Fig.2 J, O, separated sEVs were visualised by negative-positive contrast and immunogold labelling as described previously (PMID: 37103858).

Reviewer #1: Please include this justification in the revised version.

We included this justification in the revised manuscript.

(9) For Figures 3C and 3-S1, the authors should include the images used for EV quantification. Considering the concern regarding potential contamination introduced by FBS (concern 1), it is advisable for the authors to employ an independent method to identify EVs, thereby confirming the reliability of the data presented in these figures.

In our revised manuscript, we will provide all the images used for EV quantification in Figure 3C. Given that Figures 3C and 3-S1 show MV-lEVs released by HEK293T-PlamGFP cells, the possible interference by FBS-derived non-fluorescent EVs can be excluded.

Reviewer #1: Please provide all the images.

Original LASX files are provided (DOI: 10.6019/S-BIAD1456 ).

Reviewer #1: The images raising concerns regarding the contamination of EVs in FBS primarily consist of transmission electron microscopy (TEM) images, namely, Figure 1, 2P-S, S5, and Figure 4 P-U, along with the quantification of EV numbers in Figure 3. These concerns persist despite the use of fluorescent-labeled experiments. While fluorescent-labeled MV-lEVs are conclusively identified as originating from the cells, the MV-lEVs observed in Figure 1, 2P-S, S5, and Figure 4 P-U and Figure 3 may derive from both the cells and FBS.

Large EVs (with diameter >800 nm) derived from FBS were not present in our experiments, as discussed above.

(10) Do the amphiectosomes released from other cell types as well as cells in mouse kidneys or liver contain LC3B positive and CD63 positive ILVs?

Based on our confocal microscopic analysis, in addition the HEK293T-PalmGFP cells, HT29 and HepG2 cells also release similar LC3B and CD63 positive MV-lEVs. Preliminary evidence shows MV-lEV secretion by additional cell types.

The response of Reviewer #1: Please show these data in the revised manuscript. Moreover, do cells in mouse kidneys or liver contain LC3B positive and CD63 positive ILVs?

We have added new confocal microscopic images to Fig2-S3 showing amphiectosomes released also by the H9c2 (ATCC) cardiomyoblast cell line. To preserve the ultrastructure of MV-lEVs in complex organs like kidney and liver, fixation with 4% glutaraldehyde with 1% OsO4 appears to be essential. This fixation does not allow for immune detection to assess LC3B and CD63 positive MV-lEVs in the ultrathin sections.

Reviewer #2 (Public Review):

Summary:

The authors had previously identified that a colorectal cancer cell line generates small extracellular vesicles (sEVs) via a mechanism where a larger intracellular compartment containing these sEVs is secreted from the surface of the cell and then tears to release its contents. Previous studies have suggested that intraluminal vesicles (ILVs) inside endosomal multivesicular bodies and amphisomes can be secreted by the fusion of the compartment with the plasma membrane. The 'torn bag mechanism' considered in this manuscript is distinctly different because it involves initial budding off of a plasma membrane-enclosed compartment (called the amphiectosome in this manuscript, or MV-lEV). The authors successfully set out to investigate whether this mechanism is common to many cell types and to determine some of the subcellular processes involved.

The strengths of the study are:

(1) The high-quality imaging approaches used, seem to show good examples of the proposed mechanism.

(2) They screen several cell lines for these structures, also search for similar structures in vivo, and show the tearing process by real-time imaging.

(3) Regarding the intracellular mechanisms of ILV production, the authors also try to demonstrate the different stages of amphiectosome production and differently labelled ILVs using immuno-EM.

Several of these techniques are technically challenging to do well, and so these are critical strengths of the manuscript.

The weaknesses are:

(1) Most of the analysis is undertaken with cell lines. In fact, all of the analysis involving the assessment of specific proteins associated with amphiectosomes and ILVs are performed in vitro, so it is unclear whether these processes are really mirrored in vivo. The images shown in vivo only demonstrate putative amphiectosomes in the circulation, which is perhaps surprising if they normally have a short half-life and would need to pass through an endothelium to reach the vessel lumen unless they were secreted by the endothelial cells themselves.

Our previous results analyzing PFA-fixed, paraffin embedded sections of colorectal cancer patients provided direct evidence that MV-lEV secretion also occurs in humans in vivo (PMID: 31007874). Regarding your comment on the presence of amphiectosomes in the circulation despite their short half-lives, we would like to point out that Fig1.X shows a circulating lymphocyte which releases MV-lEV within the vessel lumen. Furthermore, in the revised manuscript, an additional Fig.1-S1 is provided. Here, we show the release of MV-lEVs both by an endothelial and a sub-endothelial cell (Fig.1-S1G). In addition, these images show the simultaneous presence of MV-lEVs and sEVs in the circulation (Fig.1-S1.A,C,D,H and I). The transmission electron micrographs of mouse kidney and liver sections provide additional evidence that the MV-lEVs are released by different types of cells, and the “torn bag release” also takes place in vivo (Fig.1.V).

(2) The analysis of the intracellular formation of compartments involved in the secretion process (Figure 2-S5) relies on immuno-EM, which is generally less convincing than high-/super-resolution fluorescence microscopy because the immuno-labelling is inevitably very sporadic and patchy. High-quality EM is challenging for many labs (and seems to be done very well here), but high-/super-resolution fluorescence microscopy techniques are more commonly employed, and the study already shows that these techniques should be applicable to studying the intracellular trafficking processes.

As you suggested, in the revised manuscript, we present additional super-resolution microscopy (STED) data. The intracellular formation of amphisomes, the fragmentation of LC3B-positive membranes and the formation of LC3B-positive ILVs were captured (Fig. 3B-F).

(3) One aspect of the mechanism, which needs some consideration, is what happens to the amphisome membrane, once it has budded off inside the amphiectosome. In the fluorescence images, it seems to be disrupted, but presumably, this must happen after separation from the cell to avoid the release of ILVs inside the cell. There is an additional part of Figure 1 (Figure 1Y onwards), which does not seem to be discussed in the text (and should be), that alludes to amphiectosomes often having a double membrane.

We agree with your comment regarding the amphisome membrane and we added a sentence to the Discussion of the revised manuscript. Fig1Y onwards is now discussed in the manuscript. In addition, we labelled the surface of living HEK293 cells with wheat germ agglutinin (WGA), which binds to sialic acid and N-acetyl-D-glucosamine. After removing the unbound WGA by washes, the cells were cultured for an additional 3 hours, and the release of amphiectosomes was studied. The budding amphiectosome had WGA positive membrane providing evidence that the external limiting membrane had a plasma membrane origin (Fig.3G)

(4) The real-time analysis of the amphiectosome tearing mechanism seemed relatively slow to me (over three minutes), and if this has been observed multiple times, it would be helpful to know if this is typical or whether there is considerable variation.

Thank you for this comment. In the revised manuscript, we highlight that the first released LC3 positive ILV was detected as early as within 40 sec.

Overall, I think the authors have been successful in identifying amphiectosomes secreted from multiple cell lines and demonstrating that the ILVs inside them have at least two origins (autophagosome membrane and late endosomal multivesicular body) based on the markers that they carry. The analysis of intracellular compartments producing these structures is rather less convincing and it remains unclear what cells release these structures in vivo.

I think there could be a significant impact on the EV field and consequently on our understanding of cell-cell signalling based on these findings. It will flag the importance of investigating the release of amphiectosomes in other studies, and although the authors do not discuss it, the molecular mechanisms involved in this type of 'ectosomal-style' release will be different from multivesicular compartment fusion to the plasma membrane and should be possible to be manipulated independently. Any experiments that demonstrate this would greatly strengthen the manuscript.

We appreciate these comments of the reviewer. Experiments are on their way to elucidate the mechanism of the “ectosomal style” exosome release and will be the topic of our next publication.

In general, the EV field has struggled to link up analysis of the subcellular biology of sEV secretion and the biochemical/physical analysis of the sEVs themselves, so from that perspective, the manuscript provides a novel angle on this problem.

Reviewer #3 (Public Review):

Summary:

In this manuscript, the authors describe a novel mode of release of small extracellular vesicles. These small EVs are released via the rupture of the membrane of so-called amphiectosomes that resemble "morphologically" Multivesicular Bodies.

These structures have been initially described by the authors as released by colorectal cancer cells (https://doi.org/10.1080/20013078.2019.1596668). In this manuscript, they provide experiments that allow us to generalize this process to other cells. In brief, amphiectosomes are likely released by ectocytosis of amphisomes that are formed by the fusion of multivesicular endosomes with autophagosomes. The authors propose that their model puts forward the hypothesis that LC3 positive vesicles are formed by "curling" of the autophagosomal membrane which then gives rise to an organelle where both CD63 and LC3 positive small EVs co-exist and would be released then by a budding mechanism at the cell surface that appears similar to the budding of microvesicles /ectosomes. Very correctly the authors make the distinction from migrasomes because these structures appear very similar in morphology.

Strengths:

The findings are interesting despite that it is unclear what would be the functional relevance of such a process and even how it could be induced. It points to a novel mode of release of extracellular vesicles.

Weaknesses:

This reviewer has comments and concerns concerning the interpretation of the data and the proposed model. In addition, in my opinion, some of the results in particular micrographs and immunoblots (even shown as supplementary data) are not of quality to support the conclusions.

Recommendations for the authors:

Reviewer #1 (Recommendations For The Authors):

(1) Highlight MV-IEV, ILV and limiting membrane in Figure-1G, N, and U.

Based on the suggestion, we revised Figure1

(2) Figure 1-Y-AF are not mentioned in the text.

In the revised manuscript, we discuss Figure 1Y-AF

(3) The term "IEVs" in Figure 2-S2 is not defined.

We modified the figure legend: we changed MV-lEV to amphiectosome

(4) Need to quantify co-localization in Figure 2-S2.

As suggested, we carried out the co-localisation analysis (Fig2-S2I), and Fig2-S2 was re-edited

Reviewer #2 (Recommendations For The Authors):

I have two recommendations for improving the manuscript through additional experiments:

(1) I think the description of the intracellular processes taking place in order to form amphiectosomes would be much stronger if some super-resolution imaging could be undertaken. This should label the different compartments before and after fusion with specific markers that highlight the protein signature of the different limiting and ILV membranes much more clearly than immuno-EM. It will also help in characterising the double-membrane structure of amphiectosomes at the point of budding and reveal whether the patchy labelling of the inner membrane emerges after amphiectosome release (the schematic model currently suggests that it happens before).

Thank you for your suggestion. STED microscopy was applied and results are shown in new Fig3 and the schematic model was modified accordingly.

(2) The implications of the manuscript would be more wide-ranging if the authors could test genetic manipulations that are believed to block exosome or ectosome release, eg. Rab27a or Arrdc1 knockdown. This may allow them to determine whether MV-lEVs can be released independently of the classical exosome release mechanism because they use a different route to be released from the plasma membrane. This experiment is not essential, but I think it would start to address the core regulatory mechanisms involved, and if successful, would easily allow the authors to determine the ratio of CD63-positive sEVs being secreted via classical versus amphiectosome routes.

The suggestion is very valuable for us and these studies are being performed in a separate project.

I think there are several other ways in which the manuscript could be improved to better explain some of the approaches, findings and interpretation:

(1) Include some explanation in the text of certain key tools, particularly:

a. Palm-GFP and whether its expression might alter the properties of the plasma membrane since this is used in a lot of experiments and is the only marker that seems to uniformly label the outer membrane of amphiectosomes. One concern might be that its expression drives amphiectosome secretion.

We found evidence for amphiectosome release also in the case of several different cells not expressing Palm-GFP. We believe, this excludes the possibility that Palm-GFP expression is the inducer of the amphiectosome release. Both by fluorescent and electron microscopy, the Palm-GFP non expressing cells showed very similar MV-lEVs. In addition, in the case of non-transduced HEK293 and fluorescent WGA-binding, we made similar observations.

b. Lactadherin - does this label the amphiectosomes after their release or does the wash-off step mean that it only labels cells, which subsequently release amphiectosomes?

Lactadherin labels the amphiectosomes after their release and fixation. Living cells cannot be labelled by lactadherin as PS is absent in the external plasma membrane layer of living cells. We used WGA on HEK293 cells to further support the plasma membrane origin of the external membrane of amphiectosomes.

(2) Explain the EM and confocal imaging approaches more clearly. Most importantly, is a 3D reconstruction always involved to confirm that 'separated' amphiectosomes are not joined to cells in another Z-plane.

Thank you for your suggestion. We have modified the manuscript accordingly

(3) Presenting triple-labelled images with red, green and yellow channels does not allow individual labelling to be determined without single-channel images and even then, it is much more informative to use three distinguishable colours that make a different colour with overlap, eg. CMY? Fig.2_S2D and E do not display individual channels, so definitely need to be changed.

In case of Fig.2_S2D, we now show the individual channels, the earlier E image has been removed. In case of the STED images, CMY colors had been used, as you suggested.

(4) Please discuss in the text the data in Figure 1Y onwards concerning single/double membranes on MV-lEVs.

In the revised manuscript, we discuss the question on single/double membranes and we refer to Figure 1Y-AF

(5) On line 162, reword 'intraluminal TSPAN4 only' to 'one in which TSPAN4 is only intraluminal' to make it clear that other proteins are also marking the intraluminal region, not TSPAN4 only.

We modified the text accordingly.

(6) Points for further discussion and further conclusions:

a. In vivo experiments - discuss the limitations of this part of the analysis - it seems that none of the amphiectosome markers have been analysed in this part of the study and the MV-lEVs are only in the circulation.

b. Can the authors give any further indication of the levels of MV-lEVs relative to free sEVs from any of their studies?

Using our current approach, it is not possible to determine the levels of MV-lEVs to free sEV. Without analyzing serial ultrathin sections, determination of the relative ratio of MV-lEVs and sEVs would depend on the actual section plane. In future projects, we will determine the ratio of LC3 positive and negative sEVs by single EV analysis techniques (such as SP-IRIS). In the revised manuscript, additional TEM images are included to provide evidence for the simultaneous presence of sEVs and MV-lEVs and MV-lEVs both inside and outside of the circulation.

c. Please discuss the single versus double membrane issue (relating to experiments proposed above).

We discuss this question in more details in the revised manuscript.

d. Please point out that the release mechanism (plasma membrane budding) will involve different molecular mechanisms to establish exosome release, and this might provide a route to determine relative importance.

We are currently running a systemic analysis of the release mechanism of amphiectosomes, and this will be the topic of a separate manuscript.

Reviewer #3 (Recommendations For The Authors):

* The model is not supported.

* The data is not of quality.

* The appropriate methods are not exploited.

We are sorry, we cannot respond to these unsupported critiques.

Version 2

Abstract

Abstract

Recent studies showed an unexpected complexity of extracellular vesicle (EV) biogenesis pathways. We previously found evidence that human colorectal cancer cells in vivo release large multivesicular body-like structures en bloc. Here, we tested whether this large extracellular vesicle type is unique to colorectal cancer cells. We found that all cell types we studied (including different cell lines and cells in their original tissue environment) released multivesicular large EVs (MV-lEVs). We also demonstrated that upon spontaneous rupture of the limiting membrane of the MV-lEVs, their intraluminal vesicles (ILVs) escaped to the extracellular environment by a ”torn bag mechanism”. We proved that the MV-lEVs were released by ectocytosis of amphisomes (hence, we termed them amphiectosomes). Both ILVs of amphiectosomes and small EVs separated from conditioned media were either exclusively CD63 or LC3B positive. According to our model, upon fusion of multivesicular bodies with autophagosomes, fragments of the autophagosomal inner membrane curl up to form LC3B positive ILVs of amphisomes, while CD63 positive small EVs are of multivesicular body origin. Our data suggest a novel common release mechanism for small EVs, distinct from the exocytosis of multivesicular bodies or amphisomes, as well as the small ectosome release pathway.

Introduction

Extracellular vesicles (EVs) are phospholipid bilayer enclosed structures1–4, which have important roles in cellular homeostasis and intercellular communication. Exosomes have been defined as small (~ 50-200 nm) EVs (sEVs) of endosomal origin1,3,4. Although autophagy is a major cellular homeostatic mechanism, and is implicated in a broad spectrum of human diseases, the intersection of autophagy and exosome secretion remains poorly understood. Recently, regulatory interactions have been shown between autophagy-related molecules and EV biogenesis5,6. Furthermore, the LC3-conjugation machinery was demonstrated to specify the cargo packaged into EVs7. Importantly, both others and we reported the secretion of LC3-carrying exosomes7,8. Particularly relevant to the findings presented here, is the implication of amphisomes (hybrid organelles formed by the fusion of late endosomes/multivesicular bodies (MVBs) with autophagosomes9,10) in EV biogenesis. It was suggested that fusion of the limiting membrane of amphisomes with the plasma membrane of cells results in a subsequent release of exosomes by exocytosis1,3,11. The current study was prompted by our recent data showing the in vivo en bloc release of large, MVB-like sEV clusters by human colorectal cancer cells12. Here we investigated if this was a colorectal cancer cell-specific phenomenon. Unexpectedly, we found that it was a general mechanism of sEV release that we designated as “torn bag mechanism”.

Results and Discussion

In this study, we analyzed in situ fixed, cultured cells with the released EVs preserved in their original microenvironment on a surface coated by gelatin and fibronectin. We detected large multivesicular EVs (MV-lEVs) in sections of different immersion fixed organs. We tested tumorous HT29, HepG2, and non-tumorous HEK293, HEK293T-PalmGFP, HL1 cell lines, as well as primary suspension type bone marrow derived mast cells (BMMCs). In addition, we studied ultrathin sections of mouse kidney and liver.

By the analysis of transmission electron micrographs of all tested cell types, we identified budding (Fig.1A-G) and secretion (Fig.1H-N) of MV-lEVs carrying intraluminal vesicles (ILVs). Importantly, in all cases we found evidence for the extracellular rupture of the limiting membrane of MV-lEVs and the release of ILVs (Fig.1O-U). For this novel type of sEV release, we suggest the designation “torn bag mechanism”, which is distinct from the exocytosis of MVBs and amphisomes1,3,4,11 and from the release of plasma membrane derived small EVs (sEVs) by ectocytosis13.

Transmission electron microscopic detection of the release and extracellular fate of large, multivesicular EVs secreted by different cell lines and cells in mouse organs

Major steps of the release of large, multivesicular EVs (MV-lEVs) were detected in the case of all tested cell lines including the immortal, non-tumorous HEK293T-PalmGFP (A,H,O), HEK293 (B,I,P), the tumorous cell lines HepG2 (C,J,Q) and HT29 (D,K,R), the beating cardiomyocyte cell line HL1 (E,L,S) and the primary suspension of bone marrow derived mast cells (BMMCs) (F,M,T). The different phases of EV secretion were also captured in the circulation of mouse kidney (V) and liver (W,X). According to the electron micrographs, we found evidence for the budding (A-G,X) and secretion (H-N,V,W) of the MV-lEVs. We also detected the extracellular rupture of the limiting membrane of the released MV-lEVs with the escape of the intraluminal vesicles (ILVs) by a “torn bag mechanism” (O-U,V). Although it is not always clear whether the secreted MV-lEVs have a single or double limiting membrane, several micrographs suggest the presence of the double membrane (Y-AF) in the secreted MV-lEVs. In the case of BMMCs (Y), the release phase of a multivesicular structure is captured. The bottom portion of this structure embedded in the cytoplasm is surrounded by a single membrane (white arrow head) while the upper (budding) portion is covered by double membrane (asterisk). In the shematic figures (G, N, U) the limiting membrane of MV-lEV presumably with plasma-membrane origin was indicated by red, the original limiting membrane of intracellular amphisomes, which may fragmented during the release process was indicated by blue while the ILVs of the MV-lEV were shown by gray color.

Most relevant to the in vivo conditions, we also observed the same phenomenon within the ultrathin sections of both murine kidney (Fig.1V) and liver (Fig.1W,X). In these cases, both the intact MV-lEVs (Fig.1V-X) and the “torn bag release” of sEVs (Fig.1V) were detected. Fig.1X shows that a circulating leukocyte releases MV-lEVs by ectocytosis. In Fig.1_S1, MV-lEVs, lEVs and sEVs were captured simultaneously both in kidney (Fig.1_S1C,D) and liver (Fig.1_S1F,H,I). In the mouse liver section (Fig.1_S1G), MV-lEV secretion both by endothelial and subendothelial cells can be detected.

Based on the transmission electron microscopy (TEM) analysis of ultrathin sections, it was not always obvious whether the secreted MV-lEVs had a single or double membrane. However, several micrographs suggested an at least partially intact double membrane (Fig.1Y-AF) of MV-lEVs. In the case of BMMCs (Fig.1Y), the release phase of a multivesicular structure is captured. The bottom portion of this structure, embedded in the cytoplasm, is surrounded by a single membrane while the upper (budding) portion is covered by double membrane. We hypothesize that disruption of the original amphisome membrane mainly occurs after separation of the MV-lEV from the cell to avoid the release of ILVs inside the cell.

Next, we decided to further investigate the subcellular origin of the ILVs within the secreted MV-lEVs. First, we analyzed the microenvironment of in situ fixed HEK293T-PalmGFP cells by confocal microscopy. The PalmGFP signal of HEK293T-PalmGFP cells principally associates with the plasma membrane14 (Fig.2_S1A-C), therefore the green fluorescence helped us to identify the plasma membrane-derived limiting membrane of MV-lEVs. In agreement with our previous findings on HT29 colorectal cancer cells, within the MV-lEVs, we found CD63/ALIX (Fig.2A,G), CD81/ALIX (Fig.2B,H), CD63/TSG101 (Fig.2.C,I) and CD81/TSG101 (Fig.2D,J) double positive ILVs or ILV clusters.

Detection of conventional sEV markers and the LC3 protein in HEK293T-PalmGFP cell-derived EVs

Widely used sEV markers (CD63, CD81, ALIX and TSG101) and LC3B were tested in MV-lEVs found in the microenvironment of the releasing cells by confocal microscopy after in situ fixation (A-F). Normalized fluorescence intensities were calculated in order to determine the relative localization of the limiting membrane (PalmGFP), the conventional sEV markers and the LC3B signal (G-L). Fluorescence intensity peaks of sEV markers were largely overlapping with each other, while the LC3B signal and the sEV markers showed separation. Co-localization rates were also calculated (M). The sEV markers co-localized with one another as no significant difference was found among them. In contrast, low co-localization rates were detected between the “classical” sEV markers and LC3B (one-way ANOVA, p<0.0001, n=8-26 confocal images). Real time release of LC3 positive sEVs by the “torn bag mechanism” was studied in the case of HEK293T-PalmGFP-LC3RFP cells by Elyra7 SIM2 super-resolution live cell imaging (N,O). Images were recorded continuously, and selected serial time points are shown. LC3 positive, red fluorescent small particles were released within a 5 min timeframe (O) and are indicated by white arrows. Presence of CD63 and LC3B were detected in the case of a sEV fraction separated from serum-free condition medium using immunogold transmission electronmicroscopy (TEM). HEK293T-PalmGFP derived sEV fraction is shown by negative-positive contrast without immune labelling (P). In double-labelled immunogold TEM images (Q,R), distinct LC3B positive (Q) and CD63 positive (R) sEVs were found. However, CD63-LC3B double positive EVs were not detected. Black arrowheads indicate 10 nm gold particles identifying LC3B, while white arrowheads show 5 nm gold particles corresponding to the presence of CD63. Quantitative analysis of TEM images was performed (S), and the diameters of different EV populations were determined. The LC3B negative population was significantly smaller than the LC3B positive one (p<0.0001, t-test; n=79-100). No difference was detected when the immunogold labelled sEV fraction (either LC3B positive or negative, LC3B+/-) and the unlabeled sEV fraction (sEV) were compared (p<0.05, t-test, n=112-179).

We also studied the possible autophagy related aspects of the secreted MV-lEVs. ILVs were tested for the autophagy marker LC3B in parallel with CD63 and CD81. Although LC3B, CD63 and CD81 were all present in association with the ILVs (Fig.2E,F), the LC3B and CD63 (Fig.2K) and the LC3B and CD81 (Fig.2L) signals did not overlap. Fig.2M shows that while the known sEV markers (CD63, CD81, TSG101 and ALIX) strongly co-localized with each other, LC3B positivity hardly showed colocalization with CD63 or CD81. Immunocytochemistry analysis of HT29, HepG2 and the cardiomyoblast H9c2 cells further validated the findings obtained with the HEK293T-PalmGFP cells (Fig.2_S2 and Fig.2_S3A,B). The ILVs of HEK293T-PalmGFP and HepG2 cell lines were also Rab7 positive (Fig.2_S3C,D), suggesting a late endosomal origin. Western blotting of the applied antibodies is summarized in Fig.2_S4.

The sEV markers were also tested by TEM using negative-positive contrasting technique15 on sEVs separated form serum-free conditioned medium of HEK293T-PalmGFP cells. Fig.2P confirms the typical sEV morphology. With TEM double immunogold labelling, using anti-LC3B and anti-CD63 antibodies simultaneously, we found distinct LC3B positive (Fig.2Q, Fig.2_S6O) and CD63 positive (Fig.2R, Fig.2_S6O) sEVs. Based on the analysis of TEM images, the diameters of unlabeled, and LC3B positive and negative sEVs were determined (Fig.2S). The LC3B positive sEVs had a significantly larger diameter as compared to the LC3B negative ones.

To conclude our marker studies, we detected the presence of CD63, CD81, TSG101, ALIX positive, most probably MVB-derived ILVs. In addition, the autophagosome marker carrying, LC3B positive ILVs were also found within the same single, plasma membrane limited extracellular MV-lEV, which identified these MV-lEVs as en bloc released amphisomes16 that we refer to “amphiectosomes”.

The “torn bag mechanism” was also monitored by live cell SIM2 super-resolution microscopy analysis of HEK293T-PalmGFP-LC3RFP cells (Fig.2N,O). The release of the LC3 positive red fluorescent signal was detected within a relatively short period of time (the first LC3 positive ILVs left the amphiectosome within 40s, the whole “torn bag” sEV release process was completed within 260 sec) (Fig.2O). We could rule out the possibility that rupture of the limiting membrane detected by TEM (Fig.1O-T,V) was a fixation artefact by showing the spontaneous release of LC3 positive sEVs from amphiectosomes with live-cell imaging. Characterization of the in-house developed HEK293T-PalmGFP-LC3RFP cell line is shown in Fig.2_S5.

In the following step, we addressed the question whether LC3, associated with the ILVs of MV-lEVs, indeed reflected autophagy origin. We tested MVBs (Fig.2_S6A,F,K), autophagosomes (Fig.2_S6B,G,L), amphisomes (Fig.2_S6C,H,M), amphiectosomes (Fig.2_S6D,I,N) and isolated sEV fractions of the same cells (Fig.2_S6E,J,O). Using immune electron microscopy, as expected, we found CD63 single positivity in MVBs (Fig.2_S6K). In autophagosomes, the limiting phagophore membrane was LC3B positive, and CD63 positivity was also present (Fig.2_S6L). The limiting membrane of amphisomes was LC3B negative, and the internal membranous structures were either LC3B or CD63 positive (Fig.2_S6M). The same immunoreactivity was also observed in the ILVs of the released amphiectosomes (Fig.2_S6N). Importantly, sEVs separated form serum-free conditioned medium of HEK293T-PalmGFP cells were either LC3B or CD63 positive (Fig.2_S6O). Thus, we confirmed our confocal microscopy results at the ultrastructural level. Using immunogold TEM, we provided further evidence for the budding/ectocytosis mechanism of amphiectosome release (Fig.2_S6P). The diameters of ILVs within MVBs, amphisomes and amphiectosomes were compared (Fig.2_S6Q), and the differences were likely due to the different membrane composition, pH and osmotic conditions within these structures. In agreement with our observations with separated sEVs, LC3B positive ILVs had a significantly larger diameter than the LC3B negative ones (Fig.2_S6R) possibly indicating difference in membrane composition and their different intracellular origin. Based on all the above findings, we propose the following model (Fig.3A): autophagosomes and MVBs fuse to form amphisomes, and the inner, LC3 positive membrane of autophagosomes undergoes fragmentation16. Membrane fragments curl up and form LC3 positive ILVs. Therefore, amphisomes contain both MVB-derived CD63 positive/LC3 negative and autophagosome-derived, CD63 negative/LC3 positive ILVs. The amphisome is next released from the cell by ectocytosis. Finally, the plasma membrane-derived outer membrane ruptures enabling the ILVs escape to the extracellular space by a “torn bag mechanism”. By using stimulated emission depletion (STED) microscopy, we documented the intracellular phases of our proposed model: MVB (Fig3.B), autophagosome (Fig3.D), the fusion of MVB and autophagosome (Fig3.C), fragmentation of the LC3 positive membrane and ILV formation from the membrane fragments (Fig3.F) and mature amphisome (Fig.3E). The plasma membrane origin of the external membrane of amphiectosome was further supported by wheat germ agglutinin-based live cell labelling (Fig.3G).

Amphiectosome release and its modulation

Based on our data, a model of amphiectosome release was generated (A). According to this model, the fusion of MVBs and autophagosomes forms amphisomes. The LC3B positive membrane layer (indicated in cyan) undergoes disintegration and forms LC3B positive ILVs inside the amphisome. Later, the amphisome is released into the extracellular space by ectocytosis and can be identified extracellularly as an amphiectosome. Finally, the limiting membrane(s) of the amphiectosome is ruptured and the ILVs are released as sEVs into the extracellular space by a “torn bag mechanism”. Steps of amphisome formation including LC3 positive ILV formation in 30 µM Chloroquine-treated HEK293T-PalmGFP cells was followed by super-resolution (STED) microscopy (B-F). The super-resolution STED channels were LC3B (cyan) and CD63 (magenta), while yellow indicates the confocal PalmGFP signal. Intracellular vesicular structures (such as endosomes, MVBs and amphisomes) may recieve Palm-GFP from the plasma membrane. An MVB (B), an autophagosome with Palm-GFP negative membrane (D), fusion of an autophagosome and an MVB (C), formation of LC3B positive ILVs in an amphisome (F) and a mature amphisome (E) were detected. To confirm the origin of the external membrane layer of amphiectosomes, fluorescently labelled WGA was applied. The plasma membrane of the living non-fluorescent HEK293 cells was labelled. As the external membrane of the budding amphiectosome was WGA positive, its plasma membrane origin is confirmed (G). In order to further support our model on amphiectosome release and “torn bag” EV secretion, different in vitro treatments were applied. Cytochalasin B, Colchicine, Chloroquine, Bafilomycin A1 and Rapamycin were used to modulate amphiectosome release. Targeted molecular processes are summarized (H). While Cytochalasin B inhibits actin-dependent membrane budding and cell migration, Colchicine blocks the microtubule-dependent intracellular trafficking. While Chloroquine and Bafilomycin have similar, Rapamycin has opposite effect on lysosome-autophagosome or lysosome-amphisome fusion. Chloroquine and Bafilomycin inhibit lysosomal degradation while Rapamycin accelerates it. Based on confocal microscopy, Cytochalasin B (CytoB) did not alter the dynamics of amphiectosome release (I). In contrast, both Colchicine (Colch) and Rapamycin (Rapa) significantly inhibited the release of amphiectosomes, while Chloroquine (Chloro) and Bafilomycin (Bafilo) increased the release frequency. There was no difference between the effect of Chloroquine and Bafilomycin (I). Results are shown as mean ± SD of 3-4 independent biological replicates, analyzed by one-way ANOVA and Student’s t test, *: p<0.05, **: p<0.01, ns: non-significant. Original LASX files, which served as a basis of our quantification, are publicly available (doi: 10.6019/S-BIAD1456). Example for the calculation is shown on Fig3_S1H. Presence of membrane-bound (lipidated) LC3II was tested by Western blotting. The total protein content of serum-, cell- and large EV-depleted conditioned medium of HEK293T-PalmGFP (PalmGFP) and HEK293T-PalmGFP-LC3RFP (PalmGFP-LC3RFP) cells was precipitated by TCA and 20 µg of the protein samples were loaded on the gel (J). The lipidated LC3II band was detected in all cases. Relative expression of control (Ctrl) and Chloroquine (Chloro)-treated samples were determined by densitometry. Chloroquine treatment increased the LC3II level by approximately two fold. Results are shown as mean ± SD of n=6 biological replicates.

To investigate the process of amphiectosome release, we exposed the MV-lEV releasing cells to different in vitro treatments (Fig.3H). The release of MV-lEVs was monitored by confocal microscopy of in situ fixed cell cultures. Optimal test conditions were determined (Fig.3_S1A-F) and the results are summarized in Fig.3I. Original LASX files which served as a basis of our quantification, are publicly available (doi: 10.6019/S-BIAD1456). An example for our approach of counting the MV-lEVs is shown on Fig3_S1H. Cytochalasin B did not have any effect on the discharge of MV-lEVs suggesting that the release did not involve a major actin-dependent mechanism. In contrast, there was a significant reduction of the MV-lEV secretion upon exposure of the cells to Colchicine indicating a role of microtubules in the release of the MV-lEVs. While Rapamycin significantly reduced the discharge of MV-lEVs, Chloroquine and Bafilomycin induced an enhanced MV-lEV secretion. Rapamycin activates autophagic degradation17, therefore, it induces a shift towards degradation as opposed to secretion. The lysosomotropic agents Chloroquine and Bafilomycin are known to interfere with the acidification of lysosomes18,19. By blocking the degradation pathway of MVBs/amphisomes (Fig.3H), an enhanced sEV secretion is observed. This effect is well known for exosome secretion from MVBs20,21. The diameters of the released MV-lEVs were determined based on confocal images (Fig.3_S1G). Metabolic activity of the cells was determined by a Resazurin assay, and a significant reduction was detected upon exposure of the cells to Rapamycin (Fig.3_S1D) in line with previously published data22. LC3II is the membrane-associated, lipidated autophagic form of LC323 and it is the hallmark of autophagy related membranes16. Importantly, by Western blot, we not only showed the presence of the membrane-bound LC3II in serum-free, lEV-depleted (sEV containing) conditioned medium of both HEK293T-PalmGFP and HEK293T-PalmGFP-LC3RFP cells, but the amount of LC3II substantially increased upon Chloroquine treatment (Fig.3J). Raw data of Western blots are available in Fig.3_S2.

Recent advances in the EV field shed light on migrasomes, a special type of MV-lEVs24,25. With their pomegranate-like ultrastructure, migrasomes resemble amphiectosomes. Therefore, we tested the presence of TSPAN4, a migrasome limiting membrane marker25, in amphiectosomes. Fig.4A,B,G,H show that although TSPAN4 was present intraluminally in the HEK293T-PalmGFP-derived MV-lEVs, it was clearly absent from their external membrane. Surprisingly, we identified two different HT29 cell-derived MV-lEV populations: one in which TSPAN4 was only located intraluminally (Fig.4C,E,I,K), and another one with a TSPAN4 positive external membrane (Fig.4D,F,J,L). This raised the possibility that the latter population corresponded to migrasomes. Our co-localization analysis also confirmed the existence of two distinct MV-lEV populations (Fig.4M). Next, we carried out live-cell imaging on HEK293T-PalmGFP-LC3RFP cells. The released MV-lEVs were either LC3 positive or negative intraluminally (Fig.4N,O). Our TEM images confirmed that certain cell types can release both migrasome-like structures and amphiectosomes. MV-lEVs with typical migrasome-associated retraction fiber(s) were detected in the case of HL1 (Fig.4P), HEK293T-PalmGFP (Fig.4Q) and BMMC cells (Fig.4R). Of note, it cannot be excluded that the elongated structures observed in the above cases, may correspond to tunnelling nanotubes26. Importantly, the same cell lines also released amphiectosomes by budding from the cell surface (Fig.4S-U). Taken together, based on the absence of TSPAN4 in their external membrane, and their lack of association with retraction fibers, amphiectosomes appear to be distinct from migrasomes. Besides migrasomes, another MV-lEV type was described in the case of gastrointestinal tumors and low-grade glioblastoma cells referred to as spheresome27,28. However, there is no data on a relationship of spheresome release and autophagy. Recently, endothelial cell derived, multicompartmented microvesicles (MCMVs) were shown to protrude and pinch-off from the cell surface releasing ILVs by a mechanism similar to exocytosis29. The absence of protrusion clusters described for MCMV29 distinguishes amphiectosomes from MCMVs. In addition, the previously decribed so called “nodal vesicular parcels”30 might be special examples of amphiectosomes. Finally, in C. elegans, the release autophagy and stress related large EVs (exophers) has been documented31,32,33. They contain damaged organelles and do not have a MV-lEV like ultrastructure. In contrast, the amphiectosomes we described here, have multivesicular structure without recognizable damaged organelles.

Comparison of amphiectosomes and migrasomes

Commonly used sEV markers (CD63, CD81) and TSPAN4, a suggested migrasome marker, were tested in in situ fixed intact MV-lEVs of HEK293T-PalmGFP (A,B) and HT29 (C-F) cells by confocal microscopy. Normalized fluorescence intensities were calculated in order to determine the relative localization of the limiting membrane (with PalmGFP or lactadherin labelling) and the CD63/TSPAN4 and CD81/TSPAN4 markers (G-L). In the case of HEK293T-PalmGFP-derived EVs, we did not find migrasomes with TSPAN4 in their limiting membrane. The TSPAN4 signal was only detected intraluminally in the MV-lEVs. The limiting membranes of HT29-derived MV-lEVs were either TSPAN4 positive and negative. The co-localization rate between the limiting membrane and TSPAN4 was low in case of HEK293T-PalmGFP-derived EVs. In the case of HT29 cells, two MV-lEV populations were identified: one with low and one with high co-localization rates (M). Live cell imaging of HEK293T-PalmGFP-LC3RFP cells showed retraction fiberassociated MV-lEVs with or without intraluminal LC3 positivity (N,O). Using TEM, we could identify structures with retraction fiber-associated morphology in the case of HL1 (P), HEK293T-PalmGFP (Q) and BMMC (R) cells. For comparison, budding of amphiectosomes of the same HL1 (S), HEK293TPalmGFP (T) and BMMC cells (U) are shown (without being associated with long retractions fibers).

Our approach, involving in situ fixation of cultures and tissues, made it possible to recognize sEV release from amphiectosomes by the “torn bag mechanism”. We propose that this mechanism could be easily missed earlier if conditioned medium was subjected to centrifugation, SEC purification or even to simple pipetting, which may rupture the limiting membrane of amphiectosomes. This aligns with our observation that the spontaneous escape of ILVs from untouched amphiectosomes can be completed as early as 5 minutes after amphiectosome release. Based on our data presented here, and considering that the exocytosis of MVBs/amphisomes under steady-state conditions is rarely documented in the scientific literature, we suggest that amphiectosome secretion and the “torn bag mechanism” may have a significant, yet previously unrecognized, role in sEV biogenesis.

Material and methods

Key Resources Table

Cell lines

The HEK293 human embryonic kidney, the HepG2 human hepatocyte carcinoma cell line, the HT29 human colon adenocarcinoma cell lines and the H9c2 rat cardiomyoblast cell line were purchased from the European Collection of Authenticated Cell Cultures (ECACC) through their distributor (Sigma). The HL1 cell line was purchased from Millipore. The HEK293TPalmGFP human embryonic kidney cells was kindly provided by Charles P. Lai34. Mouse bone marrow-derived mast cells (BMMCs) were differentiated and expanded as we described previously35. The HEK293, HEK293TPalmGFP, HepG2 and H9c2 cell lines were grown in DMEM (Gibco)36–38, the HT29 cells were cultured in RPMI 1640 (Gibco)12, while the HL1 cells were grown in Claycomb medium36. All cells were cultured with 10 % fetal bovine serum (FBS, BioSera) in the presence of 100 U/mL of Penicillin and 100 µg/mL Streptomycin (Sigma). Before analysis by confocal microscopy, the cells were cultured on the surface of gelatin-fibronectin coated glass coverslips (VWR). The coating solution contained 0.02 % gelatin (Sigma) and 5 mg/mL fibronectin (Invitrogen). Coverslips were coated overnight (O/N) at 37 °C.

For transmission electron microscopy (TEM), the adherent cells (HEK293, HEK293T-PalmGFP, HepG2, HT29 and HL1) were grown in gelatin-fibronectin coated 8-well Flux Cell Culture Slides (SPL).

Cell cultures were tested regularly for Mycoplasma infection by PCR, with the following PCR primers: GAAGAWATGCCWTATTTAGAAGATGG and CCRTTTTGACTYTTWCCAC-CMAGTGGTTGTTG 36.

Generation of HEK293T-PalmGFP-LC3RFP cell line

For generation of a stable HEK293T-PalmGFP-LC3RFP cell line, HEK293T-PalmGFP cells were transfected by LentiBrite RFP-LC3 Lentiviral particles (Merck) according to the instructions of the manufacturer. The GFP-RFP double positive cells were sorted by a HS800 Cell Sorter (SONY) and cell banks (master and working cell banks) were prepared. The success of the stable transfection was analyzed by immunocytochemistry and Western blotting. Results are shown in Fig.2_S5.

Confocal microscopy

Confocal microscopy was carried out as we described earlier36 with some modifications. As serum starvation significantly affects autophagy39, and EV-depleted FBS in the cell culture medium may influence cellular physiology and morphology40, FBS was not removed before fixation. Our study focuses on large EV (lEVs) with diameter > 350-500 nm. EVs in this size range, are neglectibe in FBS because of sterile filtration and heat inactivation of the serum. Unlike the majority of the studies in the field of EVs, here we analyzed untouched, in situ fixed and cultured cells together with their microenvironment. Since centrifugation may disrupt the limiting membrane of amphiectosomes, the in situ fixation made it possible to observe them in their intact form. The culture medium was gently removed by pipetting from above the cells leaving a thin medium layer only (approximately 150 µL of liquid on the cells). Without any further washing, cells were in situ fixed by 4 % paraformaldehyde (PFA) in phosphate buffer saline (PBS) for 20 min at room temperature (RT). The released lEVs were either fixed and captured during the release or were preserved on the gelatin/fibronectin surface coating. After fixation, 3x 5 min washes with 50 mM glycine in PBS were carried out. In the case of the non-fluorescent HepG2 and HT29 cells, a lactadherin-based plasma membrane staining was performed14,35,37. Lactadherin (Haematologic Technologies) was conjugated to ATTO488 fluorophore (abcam) according to the instructions of the manufacturer. The lactadherin-ATTO488 conjugate was added to the fixed cells in 1:100 dilution in PBS (for 1 h, RT) before permeabilisation. The unbound lactadherin was removed by washing with PBS (3 times, 5 min, RT) and post-fixation was carried out by 4 % PFA (20 min, RT). PFA was removed by washes with 50 mM glycine in PBS (3 times, 5 min, RT). Blocking and permeabilization of the cells were performed by 10 % FBS with 0.1 % TritonX-100 (Sigma) in PBS (1 h, RT). In general, primary antibodies were applied in 1:200 dilution overnight (O/N) at 4 °C in the above blocking and permeabilization solution. Excess primary antibodies were eliminated by washing with the blocking and permeabilization solution (3 times, 5 min RT). The secondary antibodies were applied in 1:1000 dilution in 1 % FBS in PBS (1 h, RT). Unbound secondary antibodies were eliminated by washing (1 % FBS, in PBS, 2 times, 5 min; PBS 2 times, 5 min; water 2 times, 5 min) and the samples were mounted in ProLong Diamond with DAPI (Invitrogen).

In order to provide evidence for the plasma membrane origin of the outer membrane layer of amphiectosomes, HEK293 were cultured on glass cover slips (VWR). Reaching 60 % confluency, the cells were incubated in expansion medium with 5 µg/mL CF488A conjugated Wheat Germ Agglutinin (WGA) for 30 min at 37°C. After labelling the surface of the plasma membrane by WGA, cells were washed three times by expansion medium and were cultured for an additional 3 hours. Next, they were fixed by 4 % PFA (20 min, RT). LC3B and CD63 labelling were performed as described above.

Microscopic slides were examined by Leica SP8 Lightning confocal microscope with adaptive lightning mode using a HC PL APO CS2 63x/1.40 OIL objective with hybrid detector. Where we showed released MV-lEVs, they were not joined to cells in another detected Z-plane. The applied lookup tables (LUT) were linear during this study. For image analysis and co-localization studies, we applied Leica LASX software using the unprocessed raw images. In case of co-localization studies, 20 % threshold and 10 % background settings were applied.

Multi-channel STED super-resolution imaging

Immunofluorescent labeling for multi-channel stimulated emission depletion (STED) nanoscopy was performed as in the case of confocal microscopy. The used primary antibodies were: LC3B (rabbit) and CD63 (mouse). Abberior Star 635P Goat anti Mouse Abberior Star 580 Goat anti Rabbit secondary antibodies for STED microscopy have been obtained from Abberior GmbH. Samples were mounted with SlowFade™ Diamond Antifade Mountant (Thermo). Immunofluorescence was analyzed using an Abberior Instruments Facility Line STED Microscope system built on an Olympus IX83 fully motorized inverted microscope base (Olympus), equipped with a ZDC-830 TrueFocus Z-drift compensator system, an IX3-SSU ultrasonic stage, a QUADScan Beam Scanner scanning head, APD detectors and an UPLXAPO60XO 60X oil immersion objective (NA 1.42). We used the 488, 561 and 640 nm solid state lasers for imaging, and a 775 nm solid state laser for STED depletion. Image acquisition was performed using the Imspector data acqusition software (version: 16.3.14278-w2129-win64).

Purification of sEV fraction

Small EV fractions for TEM analysis were separated from serum-free conditioned medium of HEK293T-PalmGFP cells by gravity-fitration, differencial centrifugation and tangentional flow filtration (TFF Easy, HansaBiomed) as described previosly38.

Transmission electron microscopy

Adherent cells (HEK293, HEK293T-PalmGFP, HepG2, HT29 and HL1), as well as mouse (C57BL/6) kidney and liver tissues pieces (approx. 1.5 mm x 1.5 mm) were immersed in and fixed by 4 % glutaraldehyde (48 h, 4 °C), post-fixed by 1 % osmium tetroxide (2 h, RT) and were embedded into EPON resin (Electron Microscopy Sciences) as described previously41. In the case of BMMCs, 920 μL cell suspension was complemented with 80 μL 50 % glutaraldehyde to reach the final 4 % glutaraldehyde concentration. Cells were fixed for 48 h at 4 °C and were post-fixed by 1% osmium tetroxide (2 h, RT). During sample preparation, BMMCs were collected by gravity-based sedimentation. Due to the high viscosity of EPON resin, BMMCs were embedded in LR White low viscosity resin (SPI Supplies) according to the instructions of the manufacturer. Ultrathin sections (60 nm) were contrasted by uranyl acetate (3.75 %, 10 min, RT) and lead citrate (12 min, RT).

For immunogold labelling of ultrathin sections, cells and tissues were fixed by 4 % PFA with 0.1 % glutaraldehyde (48 h, 4 °C) and were post-fixed by 0.5 % osmium tetroxide (30 min, RT). Samples were embedded into LR White hydrophilic resin. The sections were exposed to H2O2 and NaBH4 to render the epitopes accessible and were immunogold labelled as described previously12. The contrast was enhanced by uranyl acetate (3.75 %, 1 min, RT) and lead citrate (2 min, RT).

HEK293T-PalmGFP-derived sEVs separated from serum-free conditioned medium, were detected by negative-positive contrasting without embedding and sectioning15. Immunogold labelling was performed as described previously36. Antibodies were used in 1:50 dilution.

Detailed list of the used antibodies is available in the Key Resources Table.

For all electron microscopic studies, a JEOL 1011 transmission electron microscope was used. Images were captured with the help of Olympus iTEM software and for image analysis, Image J software was used.

Live cell imaging

The HEK293T-PalmGFP-LC3RFP stable cell line was cultured the same way as HEK293T-PalmGFP cells. Before the experiments, gelatin-fibronectin coated 10-well coverslip bottom chamber slide (Greiner-BioOne) was seeded and treated by 30 μM Chloroquine O/N. Release of migrasomes, amphiectosomes and sEVs were followed by the Leica SP8 Lightning confocal microscope equipped with an Okolab environmental chamber and a Zeiss ELYRA 7 with Lattice SIM² super-resolution fluorescent microscope with the help of 63x/1.4 planapochromat Oil objective. For image analysis, we applied Leica LASX, Zeiss ZEN Blue and Image J softwares.

Modulation of amphiectosome release

In order to test the release mechanism of amphiectosomes and to distinguish them from migrasomes, different treatments were applied overnight in fresh, serum containing cell culture medium except for Colchicine, where 1 h treatment was selected. Maturation and fusion of endosomes and lysosomes were inhibited by 30 μM Chloroquine (Invitrogen) or 10 nM BafilomycinA1 (Sigma). Actin polymerization was inhibited by 125 ng/mL Cytochalasin B (Sigma). Tubulin polymerization and function were inhibited by 250 pg/mL Colchicine, while an autophagy-related degradation was induced by 50 ng/mL Rapamycin. The selected concentrations were determined based both on literature data and our preliminary experiments (Fig.3_S1). Cellular metabolic activity was determined by a metabolic activity-based Resazurin assay36. Fresh cell culture medium was added to control cultures a day before the in situ fixation. Reagents were diluted in fresh cell culture medium. Leica TCS SP8 Lightning confocal microscope was used for detection of amphiectosome release. A few hundred µm2 sized area with 15-20 µm in height was tile-scanned with a few hundred cells. The MV-lEVs were recognized as CD63 positive EVs surrounded by GFP positive membrane. They were counted and were normalized to the number of nuclei. Raw images were deposited in BioImage Archive (https://www.ebi.ac.uk/bioimage-archive/) with the accession number S-BIAD1456 (doi: 10.6019/S-BIAD1456).

Western blotting

Presence of proteins and specificity of the used primary antibodies were confirmed by Western blotting as described previously36. For accurate quantification (free from variations potentially caused by EV purification), we analyzed cell-, serum- and large EV (diameter > 800 nm) free conditioned medium. The cells were cultured O/N in a serum-free culture medium. After harvesting cells were eliminated by centrifugation (300 g, 10 min at 4 °C) followed by a 2.000 g centrifugation (30 min at 4 °C) to eliminate lEVs. Total protein content of the conditioned, serum-, cell- and lEV-free medium was precipitated by trichloroacetic acid as described previously36,42. The protein pellets were suspended in cOmplete Protease Inhibitor Cocktail (Roche) containing radio-immunoprecipitation assay (RIPA) buffer.

When whole cell lysate was tested for validation of antibodies and the HEK293T-PalmGFP-LC3RFP cell line, cells were lysed in cOmplete Protease Inhibitor Cocktail (Roche) containing RIPA buffer.

Polyacrylamide gel electrophoresis was carried out using 10 % gels (acrylamide/bis-acrylamide ratio 37.5:1) or any kDa precast gels (Biorad) and a MiniProtean (BioRad) gel running system. For better solubilization of membrane proteins, equal volumes of 0.1 % TritonX-100, Laemmli buffer and samples were mixed as described previously43. Approximately 10-30 µg protein were loaded into each well. Following electrophoretic separation, proteins were transferred to PVDF membranes (Serva). Membranes were blocked with 5 % skimmed milk powder or 5 % BSA in washing buffer for 1 h. Primary antibodies were applied in 1:1000 dilution with the exception of the anti-CD63 (rabbit), anti-CD81 (rabbit) and anti-CD81 (mouse) antibodies where 1:500, 1:2500 and 1:100 dilutions were used, respectively. Peroxidase-labelled secondary antibodies were applied in 1:10 000 dilution. The signals were detected by ECL Western Blotting Substrate (Thermo Scientific) with an Imager CHEMI Premium (VWR) image analyzer system. In case of quantification, equal protein amounts were loaded to the gels. Within a biological replicate, the control and Chloroquine-treated samples were run on the same gels. To enable comparison, the relative expression of control and Chloroquine-treated samples were determined and compared.

Software and statistical analysis

For image capturing, analysis and co-localization studies, Leica LAS X, Zeiss ZEN Blue, Olympus iTEM and Image J software were used. Figures and graphs were generated using GraphPad Prism 9.4.1 and BioRender.com (Figure 1G, Figure 1N, Figure 1U, Figure 3A, Figure 3H, Figure 2 - Figure supplement 6A, Figure 2 - Figure supplement 6B, Figure 2 - Figure supplement 6C, Figure 2 - Figure supplement 6D, Figure 2 - Figure supplement 6E). For statistical analysis, standard deviation was calculated. Unpaired two tailed Student’s t-tests and one-way ANOVA were used (* p<0.05, ** p<0.01, *** p<0.001, **** p<0.0001).